---
output:
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
require(badger)
set.seed(1)
```

# tidyoperators

<!-- badges: start -->
`r badge_github_actions("tony-aw/tidyoperators")`
`r badge_repostatus("Active")`
`r badge_lifecycle("experimental")`
`r badge_custom("ORCID", "0000-0001-9498-8379", "green",   "https://orcid.org/0000-0001-9498-8379")`
<!-- badges: end -->

![](tidyoperators.svg)


The `tidyoperators` R-package adds some much needed infix operators, and a few functions, to make your R code much more tidy. It includes infix operators for the negation of logical operators (exclusive-or, not-and, not-in), safer float (in)equality operators, in-place modifying mathematical arithmetic, string arithmetic, string sub-setting, in-place modifying string arithmetic, in-place modifying string sub-setting, and in-place modifying unreal replacers. The 'tidyoperators' R-package also adds the stri_locate_ith function, which is missing from stringi. And it includes some helper functions for more complex string arithmetic. Most stringi pattern expressions options (regex, fixed, coll, charclass) are available for all string-pattern-related functions, when appropriate. This package adds the transform_if function. This package also allows integrating third-party parallel computing packages (like stringfish) for some of its functions.

WARNING: This package is still very much experimental. Function names, argument names, and so on may change dramatically. Use it for testing only, until it's stable.

CHNAGELOG (EXPERIMENTAL VERSION):

 - 8 march 2023: `stringi` is now a dependency. Completely re-written the ReadMe file, Description, and documentation.
 - 9 march 2023: added the "which"-operators.
 - 10 march 2023: `s_strapply()` now uses `stringi`, and uses `apply()` instead of `sapply()`. Renamed the which operators `%[sp]%` and `%[!sp]%` to `%[grep]%` and `%[!grep]%` to make their meaning more obvious. Added this Change log to the ReadMe file.
 - 11 march 2023: replaced the "which"-operators with the `transform_if()` function and the subset_if operators.
 - 13 march 2023: changed the name and argument convention of many of the string related functions to be more consistent. Changed the return of non-matches in the substr_-functions.

&nbsp;

## Installation

You can install `tidyoperators` from github like so:

``` r
library(devtools)
devtools::install_github("https://github.com/tony-aw/tidyoperators")
```

then load it using:

```{r}
library(tidyoperators)
```

an one can open the introduction page to the `tidyoperators` package using:

```{r eval=FALSE}
tidyoperators_help()
```


# Overview

The `tidyoperators` R package adds the following functionality:
 
 - Infix logical operators for exclusive-or, not-and, not-in, number-type, and string-type.
 - Safer (in)equality operators for floating numbers.
 - Infix operators for In-place modifiers for mathematical arithmetic.
 - Infix operators for string arithmetic.
 - Infix operators for string sub-setting.
 - The fully vectorized sub-string functions.
 - Infix operators for In-place modifying string arithmetic.
 - Infix operators for In-place modifying string sub-setting.
 - The in-place modifying unreal replacer operator.
 - `stri_locate_ith`: The stringi R package has a "locate_all" function, but no "locate_ith" function. The tidyoperators package adds the `stri_locate_ith` function, which uses the same naming and argument convention as the rest of the stringi functions, thus keeping your code consistent.
  - There are also some string helper functions, namely `s_pattern` and `s_strapply`.
 - The `transform_if` function, and some related operators.
 - The utility function `env_c`.
 - Most stringi pattern expressions options (regex, fixed, coll, charclass) are available for all string-pattern-related functions, when appropriate. 
 - This R package has only one dependency: `stringi`. No other dependencies, as to avoid "dependency hell".
 - Although this package has no other dependencies, it allows multi-threading of functions (when appropriate) through third-party packages to improve efficiency.
 

I realize there are other R-packages that cover some of the above functionalities. But I often experience that these R packages (or at least those I know of) either do not cover all that I required, had some inconsistencies, or suffered from some other significant drawbacks. Hence this package was created.

Currently this R package is only available on GitHub.

I understand one may not want to go through this entire Read-Me without knowing if the R package is worthy of your time. Therefore, allow me to give you a quick glimpse of what is possible in this R package before jumping into the details.

First, some in-place mathematical arithmetic with `tidyoperators`:

```{r}
x <- 1:10
print(x)
x %^ <-% 2 # is the same as x <- x^2
print(x)
```

Second, some numeric and string sub-types checks:

```{r}
n <- c(0:5, 0:-5, 0.1, -0.1, 0, 1, Inf, -Inf, NA, NaN)
n[n %=numtype% "B"]
n[n %=numtype% "N"]
n[n %=numtype% "I"]
n[n %=numtype% "unreal"]

s <- c(" AbcZ123 ", " abc ", " 1.3 ", " !#$%^&*() ", "  ", "  NA  ", "  NaN  ", " Inf ")
s[s %=strtype% "empty"]
s[s %=strtype% "unreal"]
s[s %=strtype% "numeric"]
s[s %=strtype% "special"]
```


And now some fun string manipulations using `tidyoperators`:

```{r}
x <- c("Hello World", "Goodbye World")

# Capitalize ONLY the ODD indices of each string:
x <- c("Hello World", "Goodbye World")
s_strapply(x, fun=\(x){
  replace(x, seq(1, length(x), 2), toupper(x)[seq(1, length(x), 2)])
})

# Extract second-last vowel of every word of every string in a vector:
x <- c("Outrageous, egregious, preposterous!", "Pleasant evening everyone")
print(x)
s_strapply(x, w=T, fun=\(x)na.omit(substr_extract(x, loc=stri_locate_ith(x, -2, regex="a|e|i|o|u", case_insensitive=TRUE))))
```

And some string arithmetic:

```{r}
"Hello" %s+% " world"
c("Hello world", "Goodbye world") %s-% " world"
c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
c("HaHa", "Ho", "Hi", "Hu", "He", "Ha") %s/% "Ha"
```

And string sub-setting:

```{r}
x <- c("yeay nay or nothing to say", "Goodmorning, goodevening and goodnight",
       paste0(letters[1:13], collapse=""))
print(x)
loc <- stri_locate_ith(x, -2, regex="ay", case_insensitive=TRUE)
substr_extract(x, loc=loc) # extract second-last "ay" in each string
substr_repl(x, "??", loc=loc) # replace second-last "ay" with "??"
substr_chartr(x, loc=loc) # switch upper/lower case of second-last "ay"
substr_addin(x, " ", "after", loc=loc) # add white space before second-last "ay"
substr_addin(x, " ", "before", loc=loc) # add white space after second-last "ay"
substr_arrange(x, "rev", loc=loc) # reverse second-last "ay"
```

If you're still interested, I invite you to read the rest of this Read-Me and perhaps try out the package yourself.

&nbsp;

# Simple additional logicals

The tidyoperators package adds a few basic logical operators:

 - `%xor%`: Exclusive OR
 - `%n&%`: NOT AND (i.e. `(!x) & (!y)`)
 - `%out%`: the opposite of `%in%` (i.e. `!x %in% y`)
 - `%?=%`: checks if both `x` and `y` are unknown or unreal (NA, NaN, Inf, -Inf)
 - `s %sgrepl% p` checks if pattern `p` (defined as either `regex`, or as a call from `s_pattern()`) appears in character vector `s` (info on the `s_pattern()` function can be found in the string section of this read-me)

Here are some examples:

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, NA, FALSE, TRUE)
y <- c(FALSE, TRUE, TRUE, FALSE, NA, NA, NA)
cbind(x, y, "x %xor% y"=x %xor% y, "x %n&% y" = x %n&% y, "x %?=% y" = x %?=% y)

1:3 %out% 1:10
1:10 %out% 1:3
```



Numbers can have many different sub-types whilst still being `numeric`. The `n %=numtype% numtype` operator will check for every value of numeric vector `n` if it can be considered a number belonging to type `numtype`. The following values for `numtype` are allowed:

 - "~0": zero, or else a number whose absolute value is smaller than the Machine tolerance (`sqrt(.Machine$double.eps)`);
 - "B": binary numbers (0 or 1);
 - "prop": proportions;
 - "N": Natural numbers (non-negative integers including zero);
 - "I": Integers;
 - "odd": odd integers;
 - "even": even integers
 - "R": Real numbers;
 - "unreal": infinity, NA, or NaN;

The string counterpart for `%=numtype%` is `s %=strtype% strtype`, which checks for every value of character vector `s` if it can seen as a certain `strtype`. The following values for `strtype` are allowed:

 - "empty": checks if the string only consists of empty spaces.
 - "unreal": checks if the string is NA, or if it has literal string "NA", "NaN" or "Inf", regardless if it has leading or trailing spaces.
 - "numeric": checks if the string can be converted to a number, disregarding leading and trailing spaces. I.e. the string "5.0" can be converted to the the actual number 5.0.
 - "special": checks if the string consists of only special characters.

Here are some examples:

```{r}

1e-20 %=numtype% "~0"
n <- c(0:5, 0:-5, 0.1, -0.1, 0, 1, Inf, -Inf, NA, NaN)
n[n %=numtype% "B"]
n[n %=numtype% "prop"]
n[n %=numtype% "B"]
n[n %=numtype% "N"]
n[n %=numtype% "I"]
n[n %=numtype% "odd"]
n[n %=numtype% "even"]
n[n %=numtype% "R"]
n[n %=numtype% "unreal"]

s <- c(" AbcZ123 ", " abc ", " 1.3 ", " !#$%^&*() ", "  ", "  NA  ", "  NaN  ", " Inf ")
s[s %=strtype% "empty"]
s[s %=strtype% "unreal"]
s[s %=strtype% "numeric"]
s[s %=strtype% "special"]
```


&nbsp;

# Safer float (in)equality operators

This package adds the `%f==%, %f!=% %f<%, %f>%, %f<=%, %f>=%` operators, which perform "float logic". They are virtually equivalent to the regular (in)equality operators,
`==, !=, <, >, <=, >=`,
except for one aspect. The float logic operators assume that if the absolute difference between `x` and `y` is smaller than the Machine tolerance, `sqrt(.Machine$double.eps)`, then `x` and `y` ought to be consider to be equal.
Thus these provide safer float (in)equality operators.
For example: `(0.1*7) == 0.7` returns `FALSE`, even though they are equal, due to the way floating numbers are stored in programming languages like R. But `(0.1*7) %f==% 0.7` returns `TRUE`.

Some examples:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)
x == y # gives FALSE, but should be TRUE
x!= y # gives TRUE, should be FALSE
x > y # not wrong
x < y # gives TRUE, should be FALSE
x %f==% y # here it's done correctly
x %f!=% y
x %f<% y # correct
x %f>% y # correct
x %f<=% y # correct
x %f>=% y # correct
```

Although designed for objects (vectors, matrices, arrays) of class `double` (floating numbers), these operators also work correctly for integers. These operators do not work for non-numeric objects.

&nbsp;

# In-place modifying mathematical arithmetic

## The problem

This R package includes infix operators for in-place modifying mathematical arithmetic. But first: what is an in-place modifier?
Consider the following line of code:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

The same expression, `mtcars$mpg[mtcars$cyl>6]`, is written twice, making this code rather long and cumbersome, even though we're just squaring the expression. The well-known `magrittr` R-package has an in-place modifier pipe, `%<>%`. This works excellent for actual functions, like so:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% mean() # same as mtcars$mpg[mtcars$cyl>6] <- mean(mtcars$mpg[mtcars$cyl>6])
```
 
But for arithmetic, one needs to translate the operators into a function, and then perform the in-place modifier pipe:


```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% raise_to_power(2)
```

This is better, but still not truly tidy.

&nbsp;

## The solution from this R package

This R package solves the above laid-out problem by implementing in-place modifying mathematical arithmetic for all mathematical operators, excluding matrix operators.

Here is a list of all in-place mathematical modifiers implemented in this R-package:

 - `x %+ <-% y` is the same as`x <- x + y`;
 - `x %- <-% y` is the same as`x <- x - y`;
 - `x %* <-% y` is the same as`x <- x * y`;
 - `x %/ <-% y` is the same as`x <- x / y`;
 - `x %^ <-% p` is the same as`x <- x^p`;
 - `x %rt <-% p` is the same as`x <- x^(1/p)`;
 - `x %logb <-% b` is the same as`x <- log(x, base=b)`;
 - `x %alogb <-% b` is the same as`x <- b^x`; if `b=exp(1)`, this is the same as`x <- exp(x)`;
 - `x %alogb <-% exp(1)` is the same as exp(x)`.


I realize there doesn't really need to be a `%logb%` operator, but since one was needed for the antilogarithm, I added the "regular" logarithm also, for consistency.
Notice that all in-place modifiers end with `   <-%` (notice the space). All infix operators that are in-place modifiers in this R package end with ` <-%`, so that it is clear what it does.

Lets look at the original problem:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

With `tidyoperators` one can now make this more tidy with the following:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %^ <-% 2
```

Much tidier, no?

&nbsp;


This is not the first or only R package that incorporates in-place modifiers. Most notably, the `inplace` R package is devoted entirely to in-place modifying mathematical arithmetic. However, `inplace` has a nasty side-effect:

If 2 R objects refer to the same values - let's say `x = 3` and `y = 3` - using an in-place modifier from the `inplace` package on `x` will also change `y`. This can be very dangerous.

**The `tidyoperators` R package does not have this problem:** modifying one object does not affect another object, even if they happen to have the same value.


&nbsp;

# Unreal replacement

Another operator added by `tidyoperators` is `x %unreal <-% y`, which replaces all NA, NaN, Inf and -Inf in `x` with the value given in `y`.

It is the same as `x[is.na(x)|is.nan(x)|is.infinite(x)] <- y`.

&nbsp;

# Missing from stringi: Locate $i^{th}$ pattern

Suppose one wants to transform all vowels in the strings of a character vector `x` such that all upper case vowels become lower case, and vice-versa.
One can do that completely in `stringi` + base R as follows:

```{r}

x <- c("HELLO WORLD", "goodbye world")
loc <- stringi::stri_locate_all(x, regex="a|e|i|o|u", case_insensitive=TRUE)
extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```

Easy peasy lemon squeezy.

&nbsp;

But now suppose one wants to transform **only** the **second-last** vowel. How are you going to do that? It's not impossible, but also not super straight-forward. For a tidy code, `stringi` really needs some kind of "stri_locate_ith" function. And, of course, the `tidyoperators` package provides just that.

The `stri_locate_ith(x, i, ...)` function locates for every element/string in character vector `x`, the $i^{th}$ occurrence of some (regex/fixed/etc) pattern. When `i` is positive, the occurrence is counted from left to right. Negative values for `i` are also allowed, in which case the occurrence is counted from the right to left. But `i=0` is not allowed though.

Thus, to get the **second** occurence of some patern, use `i=2`, and to get the **second-last** occurrence, use `i=-2`.

What this function returns exactly depends on the `simplify` argument.

If `simplify=FALSE` (the default), it returns a returns a list, one element for each string. Each list element consists of a matrix with 2 columns. The first column gives the start position of the $i^{th}$ occurence of pattern `p`. The second column gives the end position of the $i^{th}$ occurence of pattern `p`.
This list can be used in `stringi` for pattern transformation.

If `simplify=TRUE` it returns a matrix with 3 columns:

 - The first column gives the start position of the $i^{th}$ occurrence of pattern `p`.
 - The second column gives the end position of the $i^{th}$ occurrence of pattern `p`.
 - The third column gives the length of the position range of the $i^{th}$ occurrence of pattern `p`.

The `stri_locate_ith(x, i, ...)` function uses the exact same argument and naming convention as `stringi`, to keep your code consistent. And just like `stringi::stri_locate_all`, the `stri_locate_ith(x, i, ...)` function is a vectorized function: `x` and  `i` as well as the pattern (`regex, fixed, coll, charclass`) can all be different-valued vectors.

Currently, `stri_locate_ith` internally uses `mapply` for the vectorization. Once this R package is out of the `experimental` phase, I plan to replace `mapply` with a `C++` loop. But rest assured: `stri_locate_ith` is already quite fast.

&nbsp;

Now back to the original problem.

So we previosuly transformed all vowels in the strings of a character vector `x` such that all upper case vowels become lower case, and vice-versa, like so:

```{r}
x <- c("HELLO WORLD", "goodbye world")

loc <- stringi::stri_locate_all(x, regex="a|e|i|o|u", case_insensitive=TRUE)

extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```


This transforms all occurrences.

But to transform **only** the **second-last** occurence, one can now use `stri_locate_ith()` in a very similar way as was done with `stri_locate_all`:

```{r}
x <- c("HELLO WORLD", "goodbye world")

loc <- stri_locate_ith(x, -2, regex="a|e|i|o|u", case_insensitive=TRUE, simplify = FALSE) # <- this line is the key-difference

extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```

Notice that the code is virtually equivalent. We ONLY need to change the locate function. Quite tidy, right?


The transformation given in the previous code used `lapply`. I'm sure we want to stick to vectorized functions as much as possible. Therefore, the `tidyoperators` package also adds several fully vectorized `substr_`-functions, which are discussed in the very next section.

# Substr - functions

The `tidyoperators` R-package includes the following sub-string functions:

 - The `substr_repl(x, rp, ...)` function replaces a position (range) with string `rp`.
 - The `substr_chartr(x, old, new, ...)` function transforms the sub-string at a position (range) using `chartr(old, new)`. By default, it will translate upper-case characters to lower-case, and vice-versa.
 - The `substr_addin(x, addition, side, ...)` function adds the additional string `addition` at the side `side` of a position.
 - The `substr_extract(x, type, ...)` function extracts the string at, before, or after some position (range).
 - The `substr_arrange(x, arr, ...)` function sorts or reverse the sub-string at a position (range).

The "position" in the functions above can be specified either by giving the result of the `stri_locate_ith()` function (see the previous section) in argument `loc`, or by one can give manual numerical specifications using the `start, stop` or `at` arguments. 



For example:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)

loc <- stri_locate_ith(x, -2, regex="a|e|i|o|u") # find second-last occurrence of pattern

substr_extract(x, loc=loc) # extract pattern
substr_extract(x, type="before", loc=loc) # extract string before pattern
substr_extract(x, type="after", loc=loc) # extract string after pattern
substr_repl(x, "??", loc=loc) # replace pattern with double question-mark
substr_chartr(x, loc=loc) # transform case of pattern match
substr_addin(x, " ", "after", loc=loc) # add white space before pattern
substr_addin(x, " ", "before", loc=loc) # add white space after pattern

```

Simple, right?

# Basic string operations

The `tidyoperators` R package implements operators for string arithmetic and sub-setting, as well some of their in-place modifier equivalents. For consistency, and to avoid masking other common operators, all string-related operators start with `%s`, where the "s" stands for "string".

&nbsp;


## String subsetting operators

As a first sub-setting operator, we have `x %sget% ss`, which returns a subset of each string in character vector `x`. Here `ss` is a vector of length 2, or a matrix with `nrow(ss)=length(x)` and 2 columns. The object `ss` should consist entirely of non-negative integers (thus 0, 1, 2, etc. are valid, but -1, -2, -3 etc are not valid). The first element/column of ss gives the number of characters counting from the left side to be extracted from x. The second element/column of ss gives the number of characters counting from the right side to be extracted from x.

Here are 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %sget% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %sget% ss
```

Thus `%sget%` "gets" or extracts the given number of characters from the left and the right, and removes the rest.
There is also `%strim%`, which is the opposite: it trims away the number of characters from the left and right as defined in `ss`, leaving you with whatever is left.

Here are again 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %strim% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %strim% ss
```

&nbsp;


Another fun string operator is `x %ss%  s`. This essentially splits character vector `x` into a vector containing only individual characters; then this vector is subset-ted by the number given in `s`. Honestly, I do not think one would need this often, but it can be handy sometimes. For example:

```{r}
x <- "Tom Marvolo Riddle"
toupper(x) %ss% c(14, 4, 6:5, 12, 10:11, 13, 15, 8:9, 17:16, 18, 3:2, 7, 1) |>
  paste0(collapse = "")
```


&nbsp; 


## String arithmetic

Certainly, the `tidyoperators` package is not the first R package to introduce string arithmetic. But I do hope these operators have more consistent naming and functionality.

The `tidyoperators` package adds 4 arithmetic operators:

 - `x %s+% y` is the same as `paste0(x, y)`;
 - `x %s-% p` removes pattern `p` from each string in character vector `x`;
 - `x %s*% n` repeats each string in character vector `x` for `n` times;
 - `x %s/% p` counts how often pattern `p` occurs in string or vector `x`
 
I.e.:

```{r}
"Hello "%s+% " world"
c("Hello world", "Goodbye world") %s-% " world"
c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 2:7
x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
c("hello world & goodbye world", "world domination!") %s/% "world"
```

The tidyoperators R package also includes several string sub-setting functions.


It is important to note that the right-side arguments `p`, `y`, and `n` can be a single value, or a vector of the same length as `x`.

&nbsp;

## Pattern attributes in strings

The `x %s-% p` and `x %s/% p` operators (and their in-place equivalents, given later), and the `%sgrep%` operator perform pattern matching for various purposes. By default the pattern matching is interpreted as case-sensitive `regex` patterns from `stringi`.

But, of course, sometimes one wants to change this. For example, one may want it to be case insensitive. Or perhaps one wants to use fixed expressions, or something else.

The `tidyoperators` package provides options for these cases. To use more refined pattern definition, simply replace the argument/right-hand-side expression `p` in the relevant functions/operators with a call from the `s_pattern()` function.

The `s_pattern()` function uses the exact same argument convention as `stringi`. For example:

 - `s_pattern(regex=p, case_insensitive=FALSE, ...)`
 - `s_pattern(fixed=p, ...)`
 - `s_pattern(coll=p, ...)`
 - `s_pattern(boundary=p, ...)`
 - `s_pattern(charclass=p, ...)`

For consistency with base R and with packages such as `stringr`, one can also fill in `ignore.case=TRUE` or `ignore_case=TRUE` instead of `case_insensitive=TRUE`, and `s_pattern()` will still understand that.

Note that the `s_pattern()` function only works for the infix operators (i.e. functions surrounded by `%` signs).

&nbsp;

## Pattern attributes examples

```{r}
# VOWELS SUBSETTING ====

x <- c("yeay nay or nothing to say", "Goodmorning, goodevening and goodnight",
       paste0(letters[1:13], collapse=""))
print(x)
p <- s_pattern(regex = "a|e|i|o|u", case_insensitive=TRUE)

# PATTERN ARITHMETIC ====

x <- c("Hello world", "Goodbye world")
p <- s_pattern(regex=" world")
x %s-% p


x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(regex="Ha")
x %s/% p

```

And then some fixed expressions:

```{r}
x <- c("yeay yeay yeay yeay", "nay nay nay nay")
p <- s_pattern(fixed = "ye")

x <- c("Hello world", "Goodbye world")
p <- s_pattern(fixed=" world")
x %s-% p

x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(coll="Ha")
x %s/% p

```

And so on. I'm sure you get the idea.


&nbsp;

## In-place modifying string arithmetic and subsetting

With the exception of  `%ss%`, all infix operators (notice: operators, not functions) have their in-place modifying equivalent:

 - `x %s+ <-% y` is the same as `x <- x %s+% y`
 - `x %s- <-% p` is the same as `x <- x %s-% p`
 - `x %s* <-% n` is the same as `x <- x %s*% n`
 - `x %s/ <-% p` is the same as `x <- x %s/% p`
 - `x %sget <-% ss` is the same as `x <- x %sget% ss`
 - `x %strim <-% ss` is the same as `x <- x %strim% ss`


&nbsp;

# More advanced string operations with s_strapply

The string arithmetic and sub-setting operators and functions given so far - in combination with the string function already available in base R and R packages such as `stringi` - can do a lot, but it's not always flexible enough. To add extra flexibility, there is also the `s_strapply(x, fun, w=F, clp=NULL, ...)` function. This function applies the following steps to every element (every string) of character vector x:

 1) the string is split into a vector of single characters (`w=F`), or a vector of space-delimited words (`w=T`).
 2) the function `fun()` is applied to the vector from step 1.
 3) the result from step 2 is pasted together to form a single string element again, using `paste0(..., collapse=clp)`. By default, `clp=""` if `w=F`, and `clp=" "` if `w=T`.

In other words, this function turns every string in character vector `x` into its own little vector, applies a function to this vector, and pastes the returning vector together into a single string again.

This operator can be used in a very wide variety of ways.

One obvious use of this function is for re-arranging the characters or words in every string in character vector `x` in some custom way, or find the occurrence of the character on the alphabet:

```{r}
x <- c("Hello World", "Goodbye World")

# randomly shuffle letters:
s_strapply(x, sample)

# reverse word order:
s_strapply(x, rev, w=T)

# find occurrence of characters on alphabet:
s_strapply(x, fun=\(x)match(tolower(x),letters), clp="; ")

# completely customized sorting of characters (first vowels, then the rest of the letters):
custom_order <- c("a", "e", "i", "o", "u", setdiff(letters, c("a", "e", "i", "o", "u")))
print(paste0(custom_order, collapse = ""))
s_strapply(x, fun=\(x){
  rest <- setdiff(x = unique(x), y = custom_order)
  y <- factor(x = x, levels = c(custom_order, rest), ordered = TRUE)
  return(sort(y))
})
```

Much tidier and quicker, right?

Lets try something else: capitalize characters but ONLY at certain indices:

```{r}
x <- c("Hello World", "Goodbye World")
# capitalize odd indices:
s_strapply(x, fun=\(x){
  replace(x, seq(1, length(x), 2), toupper(x)[seq(1, length(x), 2)])
})

# capitalize random letters:
s_strapply(x, fun=\(x){
  ind <- sample(1:length(x), size=floor(length(x)/2))
  replace(x, ind, toupper(x[ind]))
})

# capitalize only first word:
s_strapply(x, fun=\(x){replace(x, 1, toupper(x[1]))}, w=T)
```


Lets try to take the second-last vowel of every word of every string in some character vector `x`:

```{r}
x <- c("Outrageous, egregious, preposterous!", "Pleasant evening everyone")
print(x)
p <- s_pattern(regex="a|e|i|o|u", case_insensitive = TRUE)
s_strapply(x, w=T, fun=\(x) na.omit(substr_extract(x, loc=stri_locate_ith(x, -2, regex=p))))
```

&nbsp;

# The transform_if function, and the subset_if operators

The final category of operators in this package are the subset_if operators, and the `transform_if()` function.

Consider the following code:

`x[cond(x)] <- f(x[cond(x)])`

Here a conditional subset of the object `x` is transformed with function `f`, where the condition is using a function referring to `x` itself (in this case `cond(x)`). Consequently, reference to `x` is written **four times**!
If the object has a short name like "x, this doesn't matter too much. But if the object has a longer name like `very_long_name_1`, doing something like this:

```{r eval=FALSE}
very_long_name_1[very_long_name_1 > 0] <- log(very_long_name_1[very_long_name_1 > 0])
```

becomes cumbersome, and not so tidy.

The tidyoperators package therefore adds the `transform_if()` function which will tidy this up. This function is internally defined in only vectorized functions, and without loops or apply-like functions, and is therefore quite fast.

The above code can now be re-written as:

```{r eval=FALSE}
very_long_name_1 |> transform_if(\(x)x>0, log)
```

Much tidyr, right?


Besides `transform_if`, the tidyoperators package also adds 2 "subset_if" operators:

 - The `x %[if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `TRUE`.

 - The `x %[!if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `FALSE`.

Here are a few examples to see these in action:

```{r}
object_with_very_long_name <- matrix(-10:9, ncol=2)
print(object_with_very_long_name)
object_with_very_long_name |> transform_if(\(x)x>0, log)

object_with_very_long_name %[if]% \(x)x %in% 1:10
object_with_very_long_name %[!if]% \(x)x %in% 1:10
```

Note that this function returns object `x`, to modify `x` directly, one still has to assign it. To keep your code tidy, consider combining this function with `magrittr`'s in-place modifying piper-operator (`%<>%`). I.e.:

`very_long_name_1 %<>% transform_if(cond, trans)`

&nbsp;

# Speed, efficiency, and multi-threading

This section discusses speed and optimization in the `tidyoperators` package.

## Multi-threading in string subsetting functions

All the string sub-setting functions have the `fish` argument, which is `FALSE` by default. If `fish=TRUE`, these functions will use `stringfish` functions instead of base R functions, which are faster, and also allow native multi-threading. Note that `stringfish` must be installed in order for this to work. And `stringfish` needs to be loaded also if you wish to also use multi-threading. Multi-threading in `stringfish` can be set-up by running `setoption(stringfish.nthreads=cl)` somewhere at the start of your code, where `cl` is the number of threads you want to use.

Don't use multi-threading unless you need to, as multi-threading has some overhead, thus its only faster with very large character strings.


## Multi-threading in s_strapply


The `s_strapply` function uses `apply` internally. Just like in `stri_locate_ith`, the user can multi-thread this function by replacing apply:

```{r eval=FALSE}
x <- rep(c("Hello World", "Goodbye World"), 2e5)

s_strapply(x, sort) # regular way

require(future.apply)
plan(multisession)
s_strapply(x, sort, custom_apply = future_apply) # multi-threaded way
```

Now you have a multi-threaded version of `s_strapply`.

If you combine `s_strapply` with another multi-threaded function, I advice the user to only make `s_strapply` multi-threaded, and not the function used in the `fun` argument; making multiple layers of parallel computing seems like asking for problems.

&nbsp;


# Recommended R packages

`stringi` is of course required for this packages. Besides that, I highly recommend using this R package alongside the 2 major operator-related R-packages, namely `magrittr` and `zeallot`.
&nbsp;

&nbsp;

# Compatibility with other operator-related R packages

The `stringi` R package has the `%s+%` and `%s*%` operators. They do virtually the same things as in `tidyoperators`, and so the masking of these functions can safely be ignored. I also made sure not to name any of the operators in `tidyoperators` the same as the operators in `magrittr` and `zeallot`, so that should be safe also.

&nbsp;

# Conclusion

I hope this R package will make your life a little bit tidyr.

