---
output:
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(badger)
library(knitr)
set.seed(1)
```

# tidyoperators

<!-- badges: start -->
`r badge_github_actions("tony-aw/tidyoperators")`
`r badge_repostatus("WIP")`
`r badge_lifecycle("experimental")`
`r badge_codefactor("tony-aw/tidyoperators")`
`r badge_custom("ORCID", "0000-0001-9498-8379", "green",   "https://orcid.org/0000-0001-9498-8379")`
<!-- badges: end -->

![](tidyoperators.svg)
&nbsp;

## Description

The `tidyoperators` R-package adds some much needed infix operators,
    and a few functions,
    to make your R code much more tidy.
    It includes infix operators for additional logic operators,
    safer float (in)equality operators,
    in-place modifying mathematical arithmetic,
    and infix operators for custom row- and column-wise rank-based ordering of matrices.
    It also adds some `stringi`-based string related functions, operators, and in-place modifying operators.
    This package adds the transform_if functions, and related operators.
    And finally, it also adds some functions and an operator
    for easier package/library management.
    The `tidyoperators` R-package has only one dependency, namely `stringi`,
    though it does allows multi-threading of some of the string-related functions
    (when appropriate) via the suggested `stringfish` R-package.

&nbsp;

## Changelog and status

WARNING: This package is still very much experimental. Function names, argument names, and so on may change dramatically. Use it for testing only, until it's stable.

CHANGELOG (EXPERIMENTAL VERSIONS):

 - 8 March 2023: `stringi` is now a dependency. Completely re-written the ReadMe file, Description, and documentation.
 - 9 March 2023: added the "which"-operators.
 - 10 March 2023: `s_strapply()` now uses `stringi`, and uses `apply()` instead of `sapply()`. Renamed the which operators `%[sp]%` and `%[!sp]%` to `%[grep]%` and `%[!grep]%` to make their meaning more obvious. Added this Change log to the ReadMe file.
 - 11 March 2023: replaced the "which"-operators with the `transform_if()` function and the subset_if operators.
 - 13 March 2023: changed the name and argument convention of many of the string related functions to be more consistent. Changed the return of non-matches in the substr_-functions.
 - 14 March 2023: changed the utility function to the `%m import <-%` operator. Fixed some linguistic mistakes in the documentation. Added a full documentation pdf. Fixed bugs in all `substr_`-functions. Added `codefactor` badge. Fixed some errors in the Description file. Created and added the pdf manual. Fixed some minor errors in this Read-Me markdown file.
 - 17 March 2023: added infix operators for custom row- and column-wise sorting of matrices. Slightly optimized the `substr_arrange()` function, and added the `opts_collator` argument to it. Re-ordered the sections of this Read-Me file. Adjusted the documentation the reflect the new changes.
 - 19 March 2023: removed the `s_strapply()` function in favor of the now newly added `stri_join_mat()` function and its aliases. Renamed the matrix re-order operators to `%row~%` and `%col~%`. Added a random order option to the `substr_arrange()` function. Adjusted the description, documentation, and this Read-Me file to reflect the new changes, and also fixed some spelling errors. Hopefully this will be one of the last major changes I have to make to this R package.
 - 25 March 2023: Now tests using `testthat` added for the `R CMD checks`. The `x %n&% y` operator now returns `NA` if either `x` or `y` or `NA`. The `%s+%` and `%s*%` operators now use `stringi`'s equivalent operators, for more consistency; their in-place modifiers are affected as well. Corrected some small spelling- and grammatical errors in the documentation and the Read-Me.
 - 28 March 2023: Small textual changes to the Read-Me file.
 - 4 April 2023: Added the `x % f{}% bnd` and `x % f!{}% bnd` operators. Adjusted the `%?=%` operator: now `NA %?=% Inf` and similar equality checks will also return `TRUE`. Added more tests with `testthat`. Adjusted the Read-Me file and documentations in accordance with these changes.
 - 8 April 2023: Added the `import_data()` function. Adjusted the documentation and this Read-Me file accordingly.
 - 11 April 2023: Added the `force_libPaths()` function. Shortened the Description text. Adjusted the documentation and this Read-Me file accordingly. Fixed some grammatical errors in the documentation.
 - 17 & 20 April 2023: Small textual changes to the documentation and Read-Me. file.


FUTURE PLANS:

I believe `tidyoperators` is slowly getting closer to becoming stable. There does not appear a need to add/remove many more functions/operators, although some functions, operators or arguments may need to be tweaked and/or optimized. Once I am fully satisfied with the R package (which may take a while, as I am a bit of a perfectionist), I may attempt to publish this R package to CRAN.

&nbsp;

## Installation

You can install `tidyoperators` from github like so:

``` r
remotes::install_github("https://github.com/tony-aw/tidyoperators")
```

then load it using:

```{r}
library(tidyoperators)
```

and one can open the introduction page to the `tidyoperators` package using:

```{r eval=FALSE}
tidyoperators_help()
```

&nbsp;

## Overview

The `tidyoperators` R package adds the following functionality:
 
 - Additional infix logic operators.
 - Safer (in)equality operators for float numbers.
 - Infix operators for In-place modifiers for mathematical arithmetic.
 - Infix operators for string arithmetic, string sub-setting, and their in-place modifying equivalents.
 - Infix operators for row- and column-wise rank-based re-ordering of matrices.
 - The `transform_if()` function, and some related operators.
 - The tidyoperators package adds additional `stringi` functions, namely `stri_locate_ith()` and `stri_join_mat()` (and aliases). These functions use the same naming and argument convention as the rest of the `stringi` functions, thus keeping your code consistent.
 - The fully vectorized sub-string functions, that extract, replace, add-in, transform, or re-arrange, the $i^{th}$ pattern occurrence or substring.
 - The `s_pattern()` helper function for the string infix operators.
 - The `stringi` pattern expressions options are available for all string-pattern-related functions, when appropriate.
 - Package import management operator and functions.
 - This R package has only one dependency: `stringi`. No other dependencies, as to avoid "dependency hell".
 - Although this package has no other dependencies, it allows multi-threading of functions (when appropriate) via `stringfish`.
 

Currently this R package is only available on GitHub.

I understand one may not want to go through this entire Read-Me without knowing if the R package is worthy of your time. Therefore, allow me to give you a quick glimpse of what is possible in this R package before jumping into the details.

&nbsp;

In-place mathematical arithmetic with `tidyoperators`:

```{r}
x <- 1:10
print(x)
x %^ <-% 2 # is the same as x <- x^2
print(x)
```

Simple transformations:

```{r eval=FALSE}
# in base R:
very_long_name_1[very_long_name_1 > 0] <- log(very_long_name_1[very_long_name_1 > 0])

# with tidyoperators:
very_long_name_1 %<>% transform_if(\(x)x>0, log)
```

Float equality checks:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)

x == y # gives FALSE, but should be TRUE

x %f==% y # here it's done correctly
```


Locate $i^{th}$ occurrence of some pattern in a string:

```{r}
x <- c("Goodmorning -- GOODafternoon -- GooDevening, and goodnight!",
       paste0(letters[1:13], collapse=""))
print(x)
loc <- stri_locate_ith(
  # locate second-last occurrence of "good" (ignore case) of each string in x:
  x, -2, regex="good", case_insensitive=TRUE, simplify = TRUE
)
substr(x, loc[,1], loc[,2])
```


String re-ordering using matrix re-ordering:

```{r}
# sorting words:
x <- c("Hello everyone, I'm here", "Goodbye everyone")
print(x)
mat <- stringi::stri_split_boundaries(
  x, simplify = TRUE, type="word" # vector to matrix
)
rank <- stringi::stri_rank(as.vector(mat)) |>  matrix(ncol=ncol(mat))
sorted <- mat %row~% rank # rank based matrix re-ordering
stri_c_mat(sorted, margin=1, sep=" ") # row-wise concatenate strings
```


If you're still interested, I invite you to read the rest of this Read-Me and perhaps try out the package yourself.

&nbsp;

# Additional logic operators

The tidyoperators package adds a few basic logic operators:

 - `%xor%`: Exclusive OR
 - `%n&%`: NOT AND (i.e. `(!x) & (!y)`). Note that if either `x` or `y` is `NA`, `%n&%` will also give `NA` (unlike `(!x) & (!y)`, which would give `FALSE`.)
 - `%?=%`: checks if both `x` and `y` are unknown or unreal (NA, NaN, Inf, -Inf)
 - `%out%`: the opposite of `%in%` (i.e. `!x %in% y`)
 - `s %sgrep% p` checks if pattern `p` (defined as either `regex`, or as a call from `s_pattern()`) appears in character vector `s` (info on the `s_pattern()` function can be found in the string section of this read-me)

Here are some examples:

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, NA, NaN, Inf, -Inf, TRUE, FALSE)
y <- c(FALSE, TRUE, TRUE, FALSE, rep(NA, 6))
outcome <- data.frame(
  x=x, y=y,
  "x %xor% y"=x %xor% y, "x %n&% y" = x %n&% y, "x %?=% y" = x %?=% y,
  check.names = FALSE
)
kable(outcome)

1:3 %out% 1:10
1:10 %out% 1:3
```



Numbers can have many different sub-types whilst still being `numeric`. The `n %=numtype% numtype` operator will check for every value of numeric vector `n` if it can be considered a number belonging to type `numtype`. The following values for `numtype` are allowed:

 - "~0": zero, or else a number whose absolute value is smaller than the Machine tolerance (`sqrt(.Machine$double.eps)`);
 - "B": binary numbers (0 or 1);
 - "prop": proportions;
 - "N": Natural numbers (non-negative integers including zero);
 - "I": Integers;
 - "odd": odd integers;
 - "even": even integers
 - "R": Real numbers;
 - "unreal": infinity, NA, or NaN;

The string counterpart for `%=numtype%` is `s %=strtype% strtype`, which checks for every value of character vector `s` if it can seen as a certain `strtype`. The following values for `strtype` are allowed:

 - "empty": checks if the string only consists of empty spaces.
 - "unreal": checks if the string is NA, or if it has literal string "NA", "NaN" or "Inf", regardless if it has leading or trailing spaces.
 - "numeric": checks if the string can be converted to a number, disregarding leading and trailing spaces. I.e. the string "5.0" can be converted to the the actual number 5.0.
 - "special": checks if the string consists of only special characters.

Here are some examples:

```{r}

1e-20 %=numtype% "~0"
n <- c(0:5, 0:-5, 0.1, -0.1, 0, 1, Inf, -Inf, NA, NaN)
n[n %=numtype% "B"]
n[n %=numtype% "prop"]
n[n %=numtype% "B"]
n[n %=numtype% "N"]
n[n %=numtype% "I"]
n[n %=numtype% "odd"]
n[n %=numtype% "even"]
n[n %=numtype% "R"]
n[n %=numtype% "unreal"]

s <- c(" AbcZ123 ", " abc ", " 1.3 ", " !#$%^&*() ", "  ", "  NA  ", "  NaN  ", " Inf ")
s[s %=strtype% "empty"]
s[s %=strtype% "unreal"]
s[s %=strtype% "numeric"]
s[s %=strtype% "special"]
```


&nbsp;

# Safer float (in)equality operators

This package adds the `%f==%, %f!=% %f<%, %f>%, %f<=%, %f>=%` operators, which perform "float logic". They are virtually equivalent to the regular (in)equality operators,
`==, !=, <, >, <=, >=`,
except for one aspect. The float logic operators assume that if the absolute difference between `x` and `y` is smaller than the Machine tolerance, `sqrt(.Machine$double.eps)`, then `x` and `y` ought to be consider to be equal.
Thus these provide safer float (in)equality operators.
For example: `(0.1*7) == 0.7` returns `FALSE`, even though they are equal, due to the way floating numbers are stored in programming languages like R. But `(0.1*7) %f==% 0.7` returns `TRUE`.

Some examples:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)
x == y # gives FALSE, but should be TRUE
x!= y # gives TRUE, should be FALSE
x > y # not wrong
x < y # gives TRUE, should be FALSE
x %f==% y # here it's done correctly
x %f!=% y
x %f<% y # correct
x %f>% y # correct
x %f<=% y # correct
x %f>=% y # correct

x <- c(0.3, 0.6, 0.7)
bnd <- matrix(c(x-0.1, x+0.1), ncol=2)
x %f{}% bnd
x %f!{}% bnd
```

Although designed for objects (vectors, matrices, arrays) of class `double` (floating numbers), these operators also work correctly for integers. These operators do not work for non-numeric objects.

&nbsp;

# In-place modifying mathematical arithmetic


This R package includes infix operators for in-place modifying mathematical arithmetic.

Consider the following line of code:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

The same expression, `mtcars$mpg[mtcars$cyl>6]`, is written twice, making this code rather long and cumbersome, even though we're just squaring the expression. The well-known `magrittr` R-package has an in-place modifier pipe, `%<>%`. This works excellent for actual functions, like so:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% mean() # same as mtcars$mpg[mtcars$cyl>6] <- mean(mtcars$mpg[mtcars$cyl>6])
```
 
But for arithmetic, one needs to translate the operators into a function, and then perform the in-place modifier pipe:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% raise_to_power(2)
```

This is better, but still not truly tidy.

This R package solves the above laid-out problem by implementing in-place modifying mathematical arithmetic for all mathematical operators, excluding matrix operators.

Here is a list of all in-place mathematical modifiers implemented in this R-package:

 - `x %+ <-% y` is the same as`x <- x + y`;
 - `x %- <-% y` is the same as`x <- x - y`;
 - `x %* <-% y` is the same as`x <- x * y`;
 - `x %/ <-% y` is the same as`x <- x / y`;
 - `x %^ <-% p` is the same as`x <- x^p`;
 - `x %rt <-% p` is the same as`x <- x^(1/p)`;
 - `x %logb <-% b` is the same as`x <- log(x, base=b)`;
 - `x %alogb <-% b` is the same as`x <- b^x`; if `b=exp(1)`, this is the same as`x <- exp(x)`;
 - `x %alogb <-% exp(1)` is the same as exp(x)`.

For consistency, all infix operators that are in-place modifiers in this R package end with `    <-%` (notice the space).

Lets look at the original problem:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

With `tidyoperators` one can now make this more tidy with the following:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %^ <-% 2
```

Much tidier, right?



&nbsp;

# The transform_if function, and related operators

Then we have the subset_if operators, and the `transform_if()` function.

Consider the following code:

`x[cond(x)] <- f(x[cond(x)])`

Here a conditional subset of the object `x` is transformed with function `f`, where the condition is using a function referring to `x` itself (in this case `cond(x)`). Consequently, reference to `x` is written **four times**!
If the object has a short name like "x, this doesn't matter too much. But if the object has a longer name like `very_long_name_1`, doing something like this:

```{r eval=FALSE}
very_long_name_1[very_long_name_1 > 0] <- log(very_long_name_1[very_long_name_1 > 0])
```

becomes cumbersome, and not so tidy.

The tidyoperators package therefore adds the `transform_if()` function which will tidy this up. The above code can now be re-written as:

```{r eval=FALSE}
very_long_name_1 |> transform_if(\(x)x>0, log)
```

Much tidyr, right?


Besides `transform_if`, the tidyoperators package also adds 2 "subset_if" operators:

 - The `x %[if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `TRUE`.

 - The `x %[!if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `FALSE`.

Here are a few examples to see these in action:

```{r}
object_with_very_long_name <- matrix(-10:9, ncol=2)
print(object_with_very_long_name)
object_with_very_long_name |> transform_if(\(x)x>0, log)

object_with_very_long_name %[if]% \(x)x %in% 1:10
object_with_very_long_name %[!if]% \(x)x %in% 1:10
```

Note that to modify `x` directly with `transform_if()`, one still has to assign it. To keep your code tidy, consider combining this function with `magrittr`'s in-place modifying piper-operator (`%<>%`). I.e.:

`very_long_name_1 %<>% transform_if(cond, trans)`


Another operator added by `tidyoperators` is `x %unreal <-% y`, which replaces all NA, NaN, Inf and -Inf in `x` with the value given in `y`.

So `x %unreal <-% y` is the same as `x[is.na(x)|is.nan(x)|is.infinite(x)] <- y`.


&nbsp;

# Matrix rank-based re-ordering infix operators

The `tidyoperators` R package adds 2 additional matrix operators:

 - The `x %row~% rank` operator re-orders the elements within every row of matrix `x` by the rank given in matrix `rank`.
 - The `x %col~% rank` operator re-orders the elements within every column of matrix `x` by the rank given in matrix `rank`.
 
If matrix `x` is a numeric matrix, and one wants to numerically sort the elements of every row or column, `x %row~% x` or `x %col~% x` would suffice, respectively.

If matrix `x` is not numeric, sorting using `x %row~% x` and `x %col~% x `are still possible, but probably not the best option. In the non-numeric case, providing a ranking matrix probably would be faster and give more accurate ordering.

If `rank` is a non-repeating sample of random integers (i.e. `sample(1:length(x), replace=FALSE)`), `x %row~% rank` will randomly shuffle the elements of every row, where the shuffling order of every row is independent of the other rows. Similarly, `x %col~% rank` will randomly shuffle the elements of every column, where the shuffling order of every column is independent of the other columns.

&nbsp;

Examples with a numeric matrix:

```{r}
mat <- matrix(sample(1:25), nrow=5)
print(mat)
mat %row~% mat # sort elements of every row
mat %row~% -mat # reverse-sort elements of every row
mat %col~% mat # sort elements of every column
mat %col~% -mat # reverse-sort elements of every column

mat <- matrix(sample(1:25), nrow=5)
print(mat)
rank <- sample(1:length(mat)) |> matrix(ncol=ncol(mat)) # randomized rank
mat %row~%  rank# random shuffle every row independent of other rows
mat %col~% rank # random shuffle every column independent of other columns
```


Examples with a character matrix:

```{r}
mat <- matrix(sample(letters, 25), nrow=5)
print(mat)
rank <- stringi::stri_rank(as.vector(mat)) |> matrix(ncol=ncol(mat)) # rank matrix
mat %row~% rank # sort elements of every row
mat %row~% -rank # reverse-sort elements of every row
mat %col~% rank # sort elements of every column
mat %col~% -rank # reverse-sort elements of every column

mat <- matrix(sample(letters, 25), nrow=5)
print(mat)
rank <- sample(1:length(mat)) |> matrix(ncol=ncol(mat)) # randomized rank
mat %row~%  rank# random shuffle every row independent of other rows
mat %col~% rank # random shuffle every column independent of other columns
```

These operators internally only use vectorized operations (no loops or apply-like functions), and are faster than re-ordering matrices using loops or apply-like functions.

&nbsp;

# String functions

## Matrix joining

The `tidyoperators` package adds a tiny additional function to `stringi`:

`stri_join_mat` (and their aliases `stri_c_mat` and `stri_paste_mat`).

As the name suggests, these functions perform row-wise (`margin=1`; the default) or column-wise (`margin=2`) joining of a matrix of strings, thereby transforming it to a vector of strings. You can do this already in base R, but it requires converting the matrix to a data.frame or list, and then calling `stri_join` inside `do.call()`, which to me just seems too much trouble for something *soooo* abysmally simple.

Here a practical example of their usage when re-ordering strings, words, or sentences :

```{r}
# sorting characters in strings:
x <- c("Hello world", "Goodbye world")
print(x)
mat <- stringi::stri_split_boundaries(x, simplify = TRUE, type="character")
rank <- stringi::stri_rank(as.vector(mat)) |>  matrix(ncol=ncol(mat))
sorted <- mat %row~% rank # <- fast sort matrix row-wise
print(sorted)
stri_join_mat(sorted, margin=1, sep="") # <- using new function here

# sorting words:
x <- c("Hello everyone", "Goodbye everyone")
print(x)
mat <- stringi::stri_split_boundaries(x, simplify = TRUE, type="word")
rank <- stringi::stri_rank(as.vector(mat)) |>  matrix(ncol=ncol(mat))
sorted <- mat %row~% rank # <- fast sort matrix row-wise
print(sorted)
stri_c_mat(sorted, margin=1, sep=" ") # <- alias for stri_join_mat

# randomly shuffle sentences:
x <- c("Hello, who are you? Oh, really?! Cool!", "I don't care. But I really don't.")
print(x)
mat <- stringi::stri_split_boundaries(x, simplify = TRUE, type="sentence")
rank <- sample(1:length(mat)) |> matrix(ncol=ncol(mat))
shuffled <- mat %row~% rank # <- fast sort matrix row-wise
print(shuffled)
stri_paste_mat(shuffled, margin=1, sep=" ") # <- another alias for stri_join_mat
```

&nbsp;

## Locate $i^{th}$ pattern for stringi

Suppose one wants to transform all vowels in the strings of a character vector `x` such that all upper case vowels become lower case, and vice-versa.
One can do that completely in `stringi` + base R as follows:

```{r}

x <- c("HELLO WORLD", "goodbye world")
loc <- stringi::stri_locate_all(x, regex="a|e|i|o|u", case_insensitive=TRUE)
extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```

But now suppose one wants to transform **only** the **second-last** vowel. How are you going to do that? It's not impossible, but also not super straight-forward. For a tidy code, `stringi` really needs some kind of "stri_locate_ith" function. And, of course, the `tidyoperators` package provides just that.

The `stri_locate_ith(x, i, ...)` function locates for every element/string in character vector `x`, the $i^{th}$ occurrence of some (regex/fixed/etc) pattern. When `i` is positive, the occurrence is counted from left to right. Negative values for `i` are also allowed, in which case the occurrence is counted from the right to left. But `i=0` is not allowed though. Thus, to get the **second** occurrence of some pattern, use `i=2`, and to get the **second-last** occurrence, use `i=-2`.

What this function returns exactly depends on the `simplify` argument.

If `simplify=FALSE` (the default), it returns a returns a list, one element for each string. Each list element consists of a matrix with 1 row and 2 columns. The first column gives the start position of the $i^{th}$ occurrence of a pattern. The second column gives the end position of the $i^{th}$ occurrence of a pattern.
This list can be used in `stringi` for pattern transformation.

If `simplify=TRUE` it returns a matrix with 3 columns: start position, end position, and length of position range of the $i^{th}$ occurrence of a pattern. One row for each string of character vector `x`x.

The `stri_locate_ith(x, i, ...)` function uses the exact same argument and naming convention as `stringi`, to keep your code consistent. And just like `stringi::stri_locate_all`, the `stri_locate_ith(x, i, ...)` function is a vectorized function: `x` and  `i` as well as the pattern (`regex, fixed, coll, charclass`) can all be different-valued vectors.


&nbsp;

Now back to the original problem.

So we previously transformed all vowels in the strings of a character vector `x` such that all upper case vowels become lower case, and vice-versa, like so:

```{r}
x <- c("HELLO WORLD", "goodbye world")

loc <- stringi::stri_locate_all(x, regex="a|e|i|o|u", case_insensitive=TRUE)

extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```


This transforms all occurrences.

But to transform **only** the **second-last** occurrence, one can now use `stri_locate_ith()` in a very similar way as was done with `stri_locate_all`:

```{r}
x <- c("HELLO WORLD", "goodbye world")

loc <- stri_locate_ith( # this part is the key-difference
  x, -2, regex="a|e|i|o|u", case_insensitive=TRUE, simplify = FALSE 
)

extr <- stringi::stri_sub_all(x, from=loc)
repl <- lapply(extr, \(x)chartr(x=x, old = "a-zA-Z", new = "A-Za-z"))
stringi::stri_sub_all_replace(x, loc, replacement=repl)
```

Notice that the code is virtually equivalent. We ONLY need to change the locate function.

The transformation given in the previous code used `lapply`. I'm sure we want to stick to vectorized functions as much as possible. Therefore, the `tidyoperators` package also adds several fully vectorized `substr_`-functions, which are discussed in the very next section.

&nbsp;

## Substr - functions

The `tidyoperators` R-package includes the following fully vectorized sub-string functions:

 - The `substr_repl(x, rp, ...)` function replaces a position (range) with string `rp`.
 - The `substr_chartr(x, old, new, ...)` function transforms the sub-string at a position (range) using `chartr(old, new)`. By default, it will translate upper-case characters to lower-case, and vice-versa.
 - The `substr_addin(x, addition, side, ...)` function adds the additional string `addition` at the side `side` of a position.
 - The `substr_extract(x, type, ...)` function extracts the string at, before, or after some position (range).
 - The `substr_arrange(x, arr, ...)` function re-arranges (sort, reverse, randomize) the sub-string at a position (range).

The "position" in the functions above can be specified either by giving the result of the `stri_locate_ith()` function (see the previous section) in argument `loc`, or one can give manual positions using the `start, end` or `at` arguments. 

Examples:

```{r}
x <- c("Good - good - GOOD", "Good - GOOD - good")
loc <- stri_locate_ith(
  # locate second "good" (ignore case) of each string in x:
  x, 2, regex="good", case_insensitive=TRUE 
)
```

```{r echo=FALSE}
tab <- data.frame(
  x=x,
  `substr_extract(x, loc=loc)` = substr_extract(x, loc=loc),
  `substr_extract(x, "before", loc=loc)` = substr_extract(x, "before", loc=loc),
  `substr_extract(x, "after", loc=loc)` = substr_extract(x, "after", loc=loc),
  `substr_repl(x, "??", loc=loc)` = substr_repl(x, "??", loc=loc),
  `substr_chartr(x, loc=loc)` = substr_chartr(x, loc=loc),
  `substr_addin(x, "~~", "after", loc=loc)` = substr_addin(x, "~~", "after", loc=loc),
  `substr_addin(x, "~~", "before", loc=loc)` = substr_addin(x, "~~", "before", loc=loc),
  `substr_arrange(x, loc=loc)` = substr_arrange(x, loc=loc),
  `substr_arrange(x, "decr", loc=loc)` = substr_arrange(x, "decr", loc=loc),
  `substr_arrange(x, "rev", loc=loc)` = substr_arrange(x, "rev", loc=loc),
  `substr_arrange(x, "rand", loc=loc)` = substr_arrange(x, "rand", loc=loc),
  check.names = FALSE
)
tab <- tab |> as.matrix() |> t()
kable(tab)
```


&nbsp;

# String infix operators

The `tidyoperators` R package implements infix operators for string arithmetic and sub-setting, as well some of their in-place modifier equivalents. For consistency, and to avoid masking other common operators, all string-related operators start with `%s`, where the "s" stands for "string".

&nbsp;


## String subsetting operators

As a first sub-setting operator, we have `x %sget% ss`, which returns a subset of each string in character vector `x`. Here `ss` is a vector of length 2, or a matrix with `nrow(ss)=length(x)` and 2 columns. The object `ss` should consist entirely of non-negative integers (thus 0, 1, 2, etc. are valid, but -1, -2, -3 etc are not valid). The first element/column of ss gives the number of characters counting from the left side to be extracted from x. The second element/column of ss gives the number of characters counting from the right side to be extracted from x.

Here are 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %sget% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %sget% ss
```

Thus `x %sget% ss` "gets" or extracts the given number of characters from the left and the right, and removes the rest.
There is also `x %strim% ss`, which is the opposite: it trims away the number of characters from the left and right as defined in the matrix `ss`, leaving you with whatever is left.

Here are again 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %strim% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %strim% ss
```


&nbsp; 


## String arithmetic

The `tidyoperators` package adds 4 arithmetic operators:

 - `x %s+% y` concatenates `x` and `y`;
 - `x %s-% p` removes pattern `p` from each string in character vector `x`;
 - `x %s*% n` repeats each string in character vector `x` for `n` times;
 - `x %s/% p` counts how often pattern `p` occurs in string or vector `x`
 
I.e.:

```{r}
"Hello "%s+% " world"
c("Hello world", "Goodbye world") %s-% " world"
c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 2:7
c("hello world & goodbye world", "world domination!") %s/% "world"
```

The right-side arguments `p`, `y`, and `n` can be a single value, or a vector of the same length as `x`.

&nbsp;

## Pattern attributes in strings

The `x %s-% p` and `x %s/% p` operators (and their in-place equivalents, given later), and the `%sgrep%` operator perform pattern matching for various purposes. By default the pattern matching is interpreted as case-sensitive `regex` patterns from `stringi`.

But, of course, sometimes one wants to change this. For example, one may want it to be case insensitive. Or perhaps one wants to use fixed expressions, or something else.

The `tidyoperators` package provides options for these cases. To use more refined pattern definition, simply replace the argument/right-hand-side expression `p` in the relevant operators with a call from the `s_pattern()` function.

The `s_pattern()` function uses the exact same argument convention as `stringi`. For example:

 - `s_pattern(regex=p, case_insensitive=FALSE, ...)`
 - `s_pattern(fixed=p, ...)`
 - `s_pattern(coll=p, ...)`
 - `s_pattern(boundary=p, ...)`
 - `s_pattern(charclass=p, ...)`

For consistency with base R and with packages such as `stringr`, one can also fill in `ignore.case=TRUE` or `ignore_case=TRUE` instead of `case_insensitive=TRUE`, and `s_pattern()` will still understand that.

Note that the `s_pattern()` function only works for the infix operators (i.e. operator-functions surrounded by `%` signs).

&nbsp;

## Pattern attributes examples

Regular expressions:

```{r}
x <- c("Hello world", "Goodbye world")
p <- s_pattern(regex=" world")
x %s-% p

x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(regex="Ha")
x %s/% p
```

Fixed expressions:

```{r}
x <- c("Hello world", "Goodbye world")
p <- s_pattern(fixed=" world")
x %s-% p

x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(fixed="Ha")
x %s/% p
```

And so on. I'm sure you get the idea.

&nbsp;

## In-place modifying string arithmetic and sub-setting

With the exception of  `%ss%`, all infix operators for string arithmetic and string sub-setting have their in-place modifying equivalent:

 - `x %s+ <-% y` is the same as `x <- x %s+% y`
 - `x %s- <-% p` is the same as `x <- x %s-% p`
 - `x %s* <-% n` is the same as `x <- x %s*% n`
 - `x %s/ <-% p` is the same as `x <- x %s/% p`
 - `x %sget <-% ss` is the same as `x <- x %sget% ss`
 - `x %strim <-% ss` is the same as `x <- x %strim% ss`


&nbsp;

# Package management

## force_libPaths (for simple Project Isolation)

The base R's `.libPaths()` function sets the library paths where R looks for R packages when checking or loading/attaching R packages. As such, this function could be used for proper Project Isolation. Problem, however, is that base R's `.libPaths()` function for some strange reason only allows adding new library paths, not overwrite existing system or site library paths (at least not conventionally). This stands in the way of using it for straight-forward Project Isolation.

The `tidyoperators` package therefore adds the `force_libPaths()` function, as provided in https://milesmcbain.com/posts/hacking-r-library-paths/. The `force_libPaths()` function does allow forcing R to only use the library paths exactly as specified by the user. Running this function at the start of every script in a project allows for very easy and light-weight project isolation.

Example:

```{r eval=FALSE}
force_libPaths("/mylibrary")
```


&nbsp;

## Import operator (for more Namespace control)

One can import a package and assign it to an alias in base R using:

```{r eval=FALSE}
alias <- loadNamespace("packagename")
```

Doing the above, instead of attaching a package using `library` or `require`, can (sometimes) be quite beneficial: i.e. prevent overlapping namespaces, prevent overriding base/core R functions, prevent polluting the namespace, clarify which function came from which package.

Loading a package alias does have some drawbacks. One is that you cannot easily import multiple packages under the same alias. While one would probably not want to import **multiple** packages under a single alias most of the time, there may be a couple of situations where importing multiple packages into a single alias might be actually preferable:

 - Suppose package `B` is supposed to overwrite a couple of functions from package `A` (for example if package `B` extends or improves the functionality from a function in package `A`). In that case you want to import package `A`, and then overwrite it with package `B`.
 - Although one should definitely keep the number of dependencies to a minimal, there will inevitably be packages that really need one or more dependencies. And sometimes one needs to directly access the functions from the dependencies also. Importing those dependencies under a different alias would in such a case probably be undesirable. Instead, one may want to import the dependencies under the same alias as the primary package.
 - If multiple packages kind of "belong" together, you may not want to give these packages separate aliases.

One example is the core `fastverse` + `tidyverse` data wrangling combo: `data.table` + `collapse` + `tidytable`. Considering the large amount of functions these packages have, some which unfortunately have the same name as core R functions, one would probably want to assign them in an alias object. But giving them separate aliases is perhaps undesirable: `tidytable` is supposed to overwrite some of the `data.table` functions, and one is probably always going to use these 3 packages together (i.e. they "belong" to each other).

This is where the `%m import <-%` operator comes in. It imports multiple packages under the same alias, and also informs the user which package will overwrite which function, so you will never be surprised. The `%m import <-%` operator also only loads exported functions (unlike `loadNamespace()`, which loads both internal and external functions). This is, I think, more desirable, as internal function should remain, you know, internal.

The `%m import <-%` operator will give a `warning` when the user attempts to import more than 3 packages under the same alias.

As an example, lets load `data.table`, and then `collapse`, and then `tidytable`, all under the same alias, which I will call "ftv" (for "fast-tidy-verse"):

```{r}
ftv %m import <-% c("data.table", "collapse", "tidytable")
```

Notice that the messages explain which package will overwrite what.

Now you can of course use those loaded packages as one would normally do when using a package alias.

It's better not to use this method for packages that mainly implement operators (such as this very R package), as using something like this:

```{r eval=FALSE}
to %m import <-% "tidyoperators"
to$`%row~%`(mat, rank)
```

is very cumbersome. The `%m import <-%` operator will therefore also notify the user when attempting to import an R-package where more than half of the functions are operators.

&nbsp;

## import_data

The `%m import <-%` imports everything from the package namespace. But packages often also have data sets, which are often not part of the namespace.

The `data()` function in core R can already load data from packages, but this function loads the data into the global environment, instead of returning the data directly, making assigning the data to a specific variable a bit annoying.
Therefore, the `tidyoperators` package introduces the `import_data()` function, which directly returns a data set from a package.

For example, to import the `chicago` data set from the `gamair` R package, and assign it directly to a variable (without having to do re-assignment and so on), one simply runs the following:

```{r}
d <- import_data("chicago", "gamair")
head(d)
```

That's it.

&nbsp;

## On date-based version control: the alternative to MRAN

As MRAN is no longer available, RStudio/Posit has taken up the mantle to provide date-based version control.

Simply go to https://packagemanager.rstudio.com/client/#/repos/2/overview. Then select a date and select your operating system (or, if you want to install from Source, leave it to Source). Scroll down to "Example repository setup code", and copy the code into R to use the repository.

For example, to set the repository URL in your current R script such that R packages from CRAN are downloaded and installed as they were available on 2 January 2023, when using a Microsoft Windows OS, simply run the following:

```{r}
options(repos = c(REPO_NAME = "https://packagemanager.rstudio.com/cran/2023-01-02"))
```

Using this in combination with project isolation (like using `force_libPaths()`) allows for complete control over your packages, ensuring your code will run even several years after you wrote your code.

&nbsp;

# Speed and multi-threading

All the string sub-setting functions have the `fish` argument, which is `FALSE` by default. If `fish=TRUE`, these functions will use `stringfish` functions instead of base R and `stringi` functions. The stringfish functions are usually faster, and also allow native multi-threading. Note that `stringfish` must be installed in order for this to work. And `stringfish` needs to be loaded also if you wish to also use multi-threading. Multi-threading in `stringfish` can be set-up by running `setoption(stringfish.nthreads=cl)` somewhere at the start of your code, where `cl` is the number of threads you want to use.

Don't use multi-threading unless you need to, as multi-threading has some overhead, thus its only faster with very large character vectors.

Example:

```{r}
x <- c("Goodmorning--Goodevening, and Goodnight",
       paste0(letters[1:13], collapse=""))
print(x)
loc <- stri_locate_ith(
  # locate second-last occurrence of "good" (ignore case) of each string in x:
  x, -2, regex="good", case_insensitive=TRUE 
)
substr_extract(x, loc=loc, fish = TRUE)
substr_extract(x, "before", loc=loc, fish = TRUE)
substr_extract(x, "after", loc=loc, fish = TRUE)
substr_repl(x, "??", loc=loc, fish = TRUE)
substr_chartr(x, loc=loc, fish = TRUE)
substr_addin(x, " ", "after", loc=loc, fish = TRUE) 
substr_addin(x, " ", "before", loc=loc, fish = TRUE) 
substr_arrange(x, loc=loc, fish = TRUE)
substr_arrange(x, "rev", loc=loc, fish = TRUE)
substr_arrange(x, "decr", loc=loc, fish = TRUE)
substr_arrange(x, "rand", loc=loc, fish = TRUE)

```


&nbsp;

# Recommended R packages

`stringi` is of course required for this packages. Besides that, I highly recommend using this R package alongside the 2 major operator-related R-packages, namely `magrittr` and `zeallot`.

For proper programming etiquette, I also highly recommend the following R packages:

 - The `this.path` R package (https://github.com/ArcadeAntics/this.path), which allows for getting the path of the current script (i.e. get the source file location), without all the issues found in packages such as the `here` and `whereami` packages.
 - The `rlang` R package, which one can use to get (among other things) more detailed back-traced errors, warnings, and messages.

&nbsp;

# Compatibility with other R packages

The `stringi` R package has the `%s+%` and `%s*%` operators. They do exactly the same things as in `tidyoperators`, and so the masking of these functions can safely be ignored. I also made sure not to name any of the operators in `tidyoperators` the same as the operators in `magrittr` and `zeallot`, so that should be safe also.

The `force_libPaths()` function is meant for super simple project isolation. One should probably not use the `force_libPaths()` function in a script when already using another system for project isolation like `renv` or `packrat`. It's either one or the other.

&nbsp;

# Conclusion

I hope this R package will make your life a little bit tidyr.

