---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(badger)
library(knitr)
library(tinycodet)
set.seed(1)
```

# tinycodet

<!-- badges: start -->
`r badge_github_actions("tony-aw/tinycodet")`
`r badge_repostatus("Active")`
`r badge_lifecycle("stable")`
`r badge_custom("ORCID", "0000-0001-9498-8379", "green",   "https://orcid.org/0000-0001-9498-8379")`
`r badge_custom("github-pages", "tony-aw.github.io/tinycodet", "purple",   "https://tony-aw.github.io/tinycodet")`
<!-- badges: end -->

[![](man/figures/tinycodet.svg)](https://github.com/tony-aw/tinycodet)

&nbsp;

## Description

The 'tinycodet' R-package is a tiny little R package that adds a few functions to help in your coding etiquette. It primarily focuses on 4 things:

 1) Safer decimal (in)equality testing, safer atomic conversions, and other functions for safer coding.
 2) A new package import system, that attempts to combine the benefits of using a package without attaching, with the benefits of attaching a package.
 3) Extending the string manipulation capabilities of the 'stringi' R package.
 4) Reducing repetitive code.

The 'tinycodet' R-package has only one dependency, namely 'stringi'. Most functions in this R-package are fully vectorized and optimized, and have been well documented.


```{r echo=FALSE}
rm(list=ls())
```

&nbsp;

## Installation

One can install 'tinycodet' from github like so:

```{r eval=FALSE}
remotes::install_github("https://github.com/tony-aw/tinycodet")
```

<!-- Or from CRAN (once released) like so: -->

<!-- ```{r eval=FALSE} -->
<!-- install.packages("tinycodet") -->
<!-- ``` -->

And attach the package - thus exposing its functions to the namespace - using:

```{r}
library(tinycodet)
```

And one can open the introduction help page of the 'tinycodet' package using any of the following:

```{r eval=FALSE}
?tinycodet
?tinycodet::`tinycodet-package`
?tinycodet::tinycodet_help
```



&nbsp;

## Reporting issues and giving suggestions

When you coming across an issue with the 'tinycodet' R package, you may want to report it in the "Issues" tab on the GitHub page (https://github.com/tony-aw/tinycodet/issues). If relevant, please provide reproducible R code, as that will make it easier to diagnose the issue. Please keep issue reports polite, professional, and to the point.

&nbsp;

If you have questions or ideas, please submit them in the "Discussion" tab on the GitHub page (https://github.com/tony-aw/tinycodet/discussions).


&nbsp;
