---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(badger)
library(knitr)
library(tinycodet)
set.seed(1)
```

# tinycodet


<!-- badges: start -->
`r badge_github_actions("tony-aw/tinycodet")`
`r badge_repostatus("Active")`
`r badge_lifecycle("stable")`
`r badge_custom("ORCID", "0000-0001-9498-8379", "green",   "https://orcid.org/0000-0001-9498-8379")`
`r badge_custom("github-pages", "tony-aw.github.io/tinycodet", "purple",   "https://tony-aw.github.io/tinycodet/")`
`r badge_cran_release("tinycodet")`
[![Dependencies](https://tinyverse.netlify.app/badge/tinycodet)](https://cran.r-project.org/package=tinycodet)
<!-- badges: end -->


&nbsp;

## Overview

'tinycodet' adds some functions to help in your coding etiquette. It primarily focuses on 4 aspects:

 1) Safer decimal (in)equality testing, standard-evaluated alternatives to `with()` and `aes()`, and other functions for safer coding.
 2) A new package import system, that attempts to combine the benefits of using a package without attaching, with the benefits of attaching a package.
 3) Extending the string manipulation capabilities of the 'stringi' R package.
 4) Reducing repetitive code.

'tinycodet' adheres to the [tinyverse](https://www.tinyverse.org/) philosophy (not to be confused with the tidyverse). Besides linking to 'Rcpp', 'tinycodet' has only one other dependency: 'stringi'. No other dependencies, thus avoiding "dependency hell". Most functions in this R-package are vectorized and optimised.

Although this is a relatively small R package, I do understand you may not want to go through all the articles and help files of 'tinycodet' without knowing if the R package is even worth your time. Therefore, to get a quick glimpse of what 'tinycodet' provides, I humbly invite you to take a look at the [Get Started](https://tony-aw.github.io/tinycodet/articles/tinycodet.html) page on the website (https://tony-aw.github.io/tinycodet/).



```{r echo=FALSE}
rm(list=ls())
```

&nbsp;

## Installation

One can install 'tinycodet' from CRAN like so:


```{r eval=FALSE}
install.packages("tinycodet")
```

Or get the development version from GitHub like so:

```{r eval=FALSE}
remotes::install_github("https://github.com/tony-aw/tinycodet")
```


And attach the package - thus exposing its functions to the namespace - using:

```{r}
library(tinycodet)
```

And one can open the introduction help page of the 'tinycodet' package using any of the following:

```{r eval=FALSE}
?tinycodet::tinycodet
?tinycodet::`tinycodet-package`
?tinycodet::tinycodet_help
```



&nbsp;

## Reporting issues and giving suggestions

When you coming across an issue with the 'tinycodet' R package, please report it in the "Issues" tab on the GitHub page (https://github.com/tony-aw/tinycodet/issues). If relevant, please provide reproducible R code, as that will make it easier to diagnose the issue. Please keep issue reports polite, professional, and to the point.

&nbsp;

If you have questions or ideas, please submit them in the "Discussion" tab on the GitHub page (https://github.com/tony-aw/tinycodet/discussions).

&nbsp;

