---
output:
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(badger)
library(knitr)
set.seed(1)
```

# tidyoperators

<!-- badges: start -->
`r badge_github_actions("tony-aw/tidyoperators")`
`r badge_repostatus("WIP")`
`r badge_lifecycle("experimental")`
`r badge_custom("ORCID", "0000-0001-9498-8379", "green",   "https://orcid.org/0000-0001-9498-8379")`
<!-- badges: end -->

![](tidyoperators.svg)
&nbsp;

## Description

The `tidyoperators` R-package adds some much needed infix operators,
    and a few functions,
    to make your R code much more tidy.
    It includes infix operators for additional logic operators,
    safer float (in)equality operators,
    and infix operators for custom row- and column-wise ordering of matrices.
    It also adds some `stringi`-based string related functions and operators.
    It also adds operators and a few functions to help reduce unnecessary repetitive code.
    And finally, it also adds some functions and an operator
    for easier package/library management.
    The `tidyoperators` R-package has only one dependency, namely `stringi`,
    though it does allows multi-threading of some of the string-related functions
    (when appropriate) via the suggested `stringfish` R-package.

&nbsp;

## Changelog and status

WARNING: This package is still very much experimental. Function names, argument names, and so on may change dramatically. Use it for testing only, until it's stable.

CHANGELOG (EXPERIMENTAL VERSIONS):

 - 8 March 2023: `stringi` is now a dependency. Completely re-written the ReadMe file, Description, and documentation.
 - 9 March 2023: added the "which"-operators.
 - 10 March 2023: `s_strapply()` now uses `stringi`, and uses `apply()` instead of `sapply()`. Renamed the which operators `%[sp]%` and `%[!sp]%` to `%[grep]%` and `%[!grep]%` to make their meaning more obvious. Added this Change log to the ReadMe file.
 - 11 March 2023: replaced the "which"-operators with the `transform_if()` function and the subset_if operators.
 - 13 March 2023: changed the name and argument convention of many of the string related functions to be more consistent. Changed the return of non-matches in the substr_-functions.
 - 14 March 2023: changed the utility function to the `%m import <-%` operator. Fixed some linguistic mistakes in the documentation. Added a full documentation pdf. Fixed bugs in all `substr_`-functions. Added `codefactor` badge. Fixed some errors in the Description file. Created and added the pdf manual. Fixed some minor errors in this Read-Me markdown file.
 - 17 March 2023: added infix operators for custom row- and column-wise sorting of matrices. Slightly optimized the `substr_arrange()` function, and added the `opts_collator` argument to it. Re-ordered the sections of this Read-Me file. Adjusted the documentation the reflect the new changes.
 - 19 March 2023: removed the `s_strapply()` function in favor of the now newly added `stri_join_mat()` function and its aliases. Renamed the matrix re-order operators to `%row~%` and `%col~%`. Added a random order option to the `substr_arrange()` function. Adjusted the description, documentation, and this Read-Me file to reflect the new changes, and also fixed some spelling errors. Hopefully this will be one of the last major changes I have to make to this R package.
 - 25 March 2023: Now tests using `testthat` added for the `R CMD checks`. The `x %n&% y` operator now returns `NA` if either `x` or `y` or `NA`. The `%s+%` and `%s*%` operators now use `stringi`'s equivalent operators, for more consistency; their in-place modifiers are affected as well. Corrected some small spelling- and grammatical errors in the documentation and the Read-Me.
 - 28 March 2023: Small textual changes to the Read-Me file.
 - 4 April 2023: Added the `x % f{}% bnd` and `x % f!{}% bnd` operators. Adjusted the `%?=%` operator: now `NA %?=% Inf` and similar equality checks will also return `TRUE`. Added more tests with `testthat`. Adjusted the Read-Me file and documentations in accordance with these changes.
 - 8 April 2023: Added the `import_data()` function. Adjusted the documentation and this Read-Me file accordingly.
 - 11 April 2023: Added the `force_libPaths()` function. Shortened the Description text. Adjusted the documentation and this Read-Me file accordingly. Fixed some grammatical errors in the documentation.
 - 17 & 20 April 2023: Small textual changes to the documentation and Read-Me. file.
 - 22 April 2023: Added more tests for `stri_locate_ith()`.
 - 22 May 2023: Re-arranged the sections of the Read-Me file. Removed the CodeFactor badge as it was bugging out for some reason.
 - 25 May 2023: The `stri_locate_ith()` function slightly re-written to be even faster, and improved its documentation. Edited the description. Changed the documentation regarding matrix operators, as I suspect the usage of the word "rank" might be confusing. Also edited this Read-Me file a bit.
 - 27 May 2023: Changed the naming convention of in-place modifiers to end with `   =%`. Added the `import_lsf()` function. Fixed a mistake in the documentation of the `transform_if()`.
 - 29 May 2023: The `stri_locate_ith()` now returns a matrix like `stri_locate_first/last`. Moreover, I replaced the `mapply` call with only vectorized functions; `stri_locate_ith()` is now almost as fast as the `stringi` functions it calls.
 - 30 May 2023: Fixed some lingering mistakes in the documentations, that were left over since the changes from 29 May 2023.
 - 4 June 2023: Major changes this time. The `tranform_if()` function now allows the user to specify 2 transformation functions, one for if the condition is `TRUE`, and the other for `FALSE`. Added the `import_inops()` function. Changed the `alias %m import <-% pkgs` operator into the `import_as()` function. Added the `pkgs %installed in% lib.loc%` operator. Re-arranged this Read-Me file a bit.


FUTURE PLANS:

I believe `tidyoperators` is slowly getting closer to becoming stable. There does not appear a need to add/remove many more functions/operators, although some functions, operators or arguments may need to be tweaked and/or optimized. Once I am fully satisfied with the R package (which may take a while, as I am a bit of a perfectionist), I may attempt to publish this R package to CRAN.


&nbsp;


## Installation

You can install `tidyoperators` from github like so:

``` r
remotes::install_github("https://github.com/tony-aw/tidyoperators")
```

then load it using:

```{r}
library(tidyoperators)
```

and one can open the introduction page to the `tidyoperators` package using:

```{r eval=FALSE}
tidyoperators_help()
```

&nbsp;

## Overview

The `tidyoperators` R package adds the following functionality:
 
 - Additional infix logic operators.
 - Safer (in)equality operators for float numbers.
 - Infix operators for row- and column-wise re-ordering of matrices.
 - Several in-place modifying arithmetic and string operators, and a few functions, to help reduce repetitive codes, for the sake of the "Don't Repeat Yourself"-coding principle (`DRY`).
 - Package import management operator and functions.
 - The tidyoperators package adds additional `stringi` functions, namely `stri_locate_ith()` and `stri_join_mat()` (and aliases). These functions use the same naming and argument convention as the rest of the `stringi` functions, thus keeping your code consistent.
 - The fully vectorized sub-string functions, that extract, replace, add-in, transform, or re-arrange, the $i^\textrm{th}$ pattern occurrence or substring.
 - The `s_pattern()` helper function for the string infix operators.
 - The `stringi` pattern expressions options are available for all string-pattern-related functions, when appropriate.
 - This R package has only one dependency: `stringi`. No other dependencies, as to avoid "dependency hell".
 - Although this package has no other dependencies, it allows multi-threading of functions (when appropriate) via `stringfish`.


Currently this R package is only available on GitHub.

I understand one may not want to go through this entire Read-Me without knowing if the R package is worthy of your time. Therefore, allow me to give you a quick glimpse of what is possible in this R package before jumping into the details.

&nbsp;

Operators and functions for the "Don't Repeat Yourself" principle:

```{r eval=FALSE}
# in base R:
very_long_name_1[very_long_name_1 > 0] <- log(very_long_name_1[very_long_name_1 > 0])
very_long_name_1 <- very_long_name_1^2

# with tidyoperators:
very_long_name_1 %<>% transform_if(\(x)x>0, log)
very_long_name_1 %^ =% 2
```

Safer float equality checks:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)

x == y # gives FALSE, but should be TRUE

x %f==% y # here it's done correctly
```


Locate $i^\textrm{th}$ occurrence of some pattern in a string:

```{r}
x <- c("Goodmorning -- GOODafternoon -- GooDevening, and goodnight!",
       paste0(letters[1:13], collapse=""))
print(x)
loc <- stri_locate_ith(
  # locate second-last occurrence of "good" (ignore case) of each string in x:
  x, -2, regex="good", case_insensitive=TRUE)
substr(x, loc[,1], loc[,2])
```


String re-ordering using matrix re-ordering:

```{r}
# sorting words:
x <- c("Hello everyone, I'm here", "Goodbye everyone")
print(x)
x <- stringi::stri_split_boundaries(
  x, simplify = TRUE, type="word" # vector to matrix
)
mat <- stringi::stri_rank(as.vector(x)) |>
  matrix(ncol=ncol(x)) # matrix of ordering ranks
sorted <- x %row~% mat # matrix re-ordering
stri_c_mat(sorted, margin=1, sep=" ") # row-wise concatenate strings
```


If you're still interested, I invite you to read the rest of this Read-Me and perhaps try out the package yourself.

&nbsp;

# Additional logic operators

The tidyoperators package adds a few basic logic operators:

 - `%xor%`: Exclusive OR
 - `%n&%`: NOT AND (i.e. `(!x) & (!y)`). Note that if either `x` or `y` is `NA`, `%n&%` will also give `NA` (unlike `(!x) & (!y)`, which would give `FALSE`.)
 - `%?=%`: checks if both `x` and `y` are unknown or unreal (NA, NaN, Inf, -Inf)
 - `%out%`: the opposite of `%in%` (i.e. `!x %in% y`)
 - `s %sgrep% p` checks if pattern `p` (defined as either `regex`, or as a call from `s_pattern()`) appears in character vector `s` (info on the `s_pattern()` function can be found in the string section of this read-me)

Here are some examples:

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, NA, NaN, Inf, -Inf, TRUE, FALSE)
y <- c(FALSE, TRUE, TRUE, FALSE, rep(NA, 6))
outcome <- data.frame(
  x=x, y=y,
  "x %xor% y"=x %xor% y, "x %n&% y" = x %n&% y, "x %?=% y" = x %?=% y,
  check.names = FALSE
)
kable(outcome)

1:3 %out% 1:10
1:10 %out% 1:3
```



Numbers can have many different sub-types whilst still being `numeric`. The `n %=numtype% numtype` operator will check for every value of numeric vector `n` if it can be considered a number belonging to type `numtype`. The following values for `numtype` are allowed:

 - "~0": zero, or else a number whose absolute value is smaller than the Machine tolerance (`sqrt(.Machine$double.eps)`);
 - "B": binary numbers (0 or 1);
 - "prop": proportions;
 - "N": Natural numbers (non-negative integers including zero);
 - "I": Integers;
 - "odd": odd integers;
 - "even": even integers
 - "R": Real numbers;
 - "unreal": infinity, NA, or NaN;

The string counterpart for `%=numtype%` is `s %=strtype% strtype`, which checks for every value of character vector `s` if it can seen as a certain `strtype`. The following values for `strtype` are allowed:

 - "empty": checks if the string only consists of empty spaces.
 - "unreal": checks if the string is NA, or if it has literal string "NA", "NaN" or "Inf", regardless if it has leading or trailing spaces.
 - "numeric": checks if the string can be converted to a number, disregarding leading and trailing spaces. I.e. the string "5.0" can be converted to the the actual number 5.0.
 - "special": checks if the string consists of only special characters.

Here are some examples:

```{r}

1e-20 %=numtype% "~0"
n <- c(0:5, 0:-5, 0.1, -0.1, 0, 1, Inf, -Inf, NA, NaN)
n[n %=numtype% "B"]
n[n %=numtype% "prop"]
n[n %=numtype% "B"]
n[n %=numtype% "N"]
n[n %=numtype% "I"]
n[n %=numtype% "odd"]
n[n %=numtype% "even"]
n[n %=numtype% "R"]
n[n %=numtype% "unreal"]

s <- c(" AbcZ123 ", " abc ", " 1.3 ", " !#$%^&*() ", "  ", "  NA  ", "  NaN  ", " Inf ")
s[s %=strtype% "empty"]
s[s %=strtype% "unreal"]
s[s %=strtype% "numeric"]
s[s %=strtype% "special"]
```


&nbsp;

# Safer float (in)equality operators

This package adds the `%f==%, %f!=% %f<%, %f>%, %f<=%, %f>=%` operators, which perform a form of "float logic". They are virtually equivalent to the regular (in)equality operators,
`==, !=, <, >, <=, >=`,
except for one aspect. The float logic operators assume that if the absolute difference between `x` and `y` is smaller than the Machine tolerance, `sqrt(.Machine$double.eps)`, then `x` and `y` ought to be consider to be equal.
Thus these provide safer float (in)equality operators.
For example: `(0.1*7) == 0.7` returns `FALSE`, even though they are equal, due to the way floating numbers are stored in programming languages like R, Python, etc. But `(0.1*7) %f==% 0.7` returns `TRUE`.

Some examples:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)
x == y # gives FALSE, but should be TRUE
x!= y # gives TRUE, should be FALSE
x > y # not wrong
x < y # gives TRUE, should be FALSE
x %f==% y # here it's done correctly
x %f!=% y
x %f<% y # correct
x %f>% y # correct
x %f<=% y # correct
x %f>=% y # correct

x <- c(0.3, 0.6, 0.7)
bnd <- matrix(c(x-0.1, x+0.1), ncol=2)
x %f{}% bnd
x %f!{}% bnd
```

Although designed for objects (vectors, matrices, arrays) of class `double` (floating numbers), these operators also work correctly for integers. These operators do not work for non-numeric objects.


&nbsp;

# Matrix re-ordering operators

The `tidyoperators` R package adds 2 additional matrix operators:

 - The `x %row~% mat` operator re-orders the elements within every row of matrix `x` by the ordering ranks given in matrix `mat`.
 - The `x %col~% mat` operator re-orders the elements within every column of matrix `x` by the ordering ranks given in matrix `mat`.
 
If matrix `x` is a numeric matrix, and one wants to numerically sort the elements of every row or column, `x %row~% x` or `x %col~% x` would suffice, respectively.

If matrix `x` is not numeric, sorting using `x %row~% x` and `x %col~% x `are still possible, but probably not the best option. In the non-numeric case, providing a matrix of ordering ranks would probably be faster and give more accurate ordering.

If `mat` is a matrix of non-repeating random integers (i.e. `sample(1:length(x), replace=FALSE)`), `x %row~% mat` will randomly shuffle the elements of every row, where the shuffling order of every row is independent of the other rows. Similarly, `x %col~% mat` will randomly shuffle the elements of every column, where the shuffling order of every column is independent of the other columns.

&nbsp;

Examples with a numeric matrix:

```{r}
x <- matrix(sample(1:25), nrow=5)
print(x)
x %row~% x # sort elements of every row
x %row~% -x # reverse-sort elements of every row
x %col~% x # sort elements of every column
x %col~% -x # reverse-sort elements of every column

x <- matrix(sample(1:25), nrow=5)
print(x)
rand <- sample(1:length(x)) |> matrix(ncol=ncol(x)) # matrix of random integers
x %row~% rand # random shuffle every row independent of other rows
x %col~% rand # random shuffle every column independent of other columns
```


Examples with a character matrix:

```{r}
x <- matrix(sample(letters, 25), nrow=5)
print(x)
mat <- stringi::stri_rank(as.vector(x)) |>
  matrix(ncol=ncol(x)) # matrix of ordering ranks
x %row~% mat # sort elements of every row
x %row~% -mat # reverse-sort elements of every row
x %col~% mat # sort elements of every column
x %col~% -mat # reverse-sort elements of every column

x <- matrix(sample(letters, 25), nrow=5)
print(x)
rand <- sample(1:length(x)) |> matrix(ncol=ncol(x)) # matrix of random integers
x %row~% rand # random shuffle every row independent of other rows
x %col~% rand # random shuffle every column independent of other columns
```

These operators internally only use vectorized operations (no loops or apply-like functions), and are faster than re-ordering matrices using loops or apply-like functions.

&nbsp;

# String functions

## Matrix joining

The `tidyoperators` package adds a tiny additional function to `stringi`:

`stri_join_mat` (and their aliases `stri_c_mat` and `stri_paste_mat`).

As the name suggests, these functions perform row-wise (`margin=1`; the default) or column-wise (`margin=2`) joining of a matrix of strings, thereby transforming it to a vector of strings. You can do this already in base R, but it requires converting the matrix to a data.frame or list, and then calling `stri_join` inside `do.call()`, which to me just seems too much trouble for something *soooo* abysmally simple.

Here is an example of their usage when re-ordering strings, words, or sentences :

```{r}
# sorting characters in strings:
x <- c("Hello world", "Goodbye world")
print(x)
x <- stringi::stri_split_boundaries(x, simplify = TRUE, type="character")
mat <- stringi::stri_rank(as.vector(x)) |>  matrix(ncol=ncol(x))
sorted <- x %row~% mat # fast sort matrix row-wise
stri_join_mat(sorted, margin=1, sep="") # <- using new function here

# sorting words:
x <- c("Hello everyone", "Goodbye everyone")
print(x)
x <- stringi::stri_split_boundaries(x, simplify = TRUE, type="word")
mat <- stringi::stri_rank(as.vector(x)) |>  matrix(ncol=ncol(x))
sorted <- x %row~% mat # <- fast sort matrix row-wise
stri_c_mat(sorted, margin=1, sep=" ") # <- alias for stri_join_mat

# randomly shuffle sentences:
x <- c("Hello, who are you? Oh, really?! Cool!", "I don't care. But I really don't.")
print(x)
x <- stringi::stri_split_boundaries(x, simplify = TRUE, type="sentence")
mat <- sample(1:length(x)) |> matrix(ncol=ncol(x))
shuffled <- x %row~% mat # <- fast sort matrix row-wise
print(shuffled)
stri_paste_mat(shuffled, margin=1, sep=" ") # <- another alias for stri_join_mat
```

&nbsp;

## Locate $i^\textrm{th}$ pattern for stringi

Suppose one wants to transform the **first** vowels in the strings of a character vector `str`, such that all upper case vowels become lower case, and vice-versa.
One can do that completely in `stringi` + base R as follows:

```{r}

x <- c("HELLO WORLD", "goodbye world")
loc <- stringi::stri_locate_first(x, regex="a|e|i|o|u", case_insensitive=TRUE)
extr <- stringi::stri_sub(x, from=loc)
repl <- chartr(extr, old = "a-zA-Z", new = "A-Za-z")
stringi::stri_sub_replace(x, loc, replacement=repl)
```

But now suppose one wants to transform the **second-last** vowel. How are you going to do that? It's not impossible, but also not super straight-forward. For a tidy code, `stringi` really needs some kind of "stri_locate_ith" function. And, of course, the `tidyoperators` package provides just that.

The `stri_locate_ith(str, i, ...)` function locates for every element/string in character vector `str`, the $i^\textrm{th}$ occurrence of some (regex/fixed/etc) pattern. When `i` is positive, the occurrence is counted from left to right. Negative values for `i` are also allowed, in which case the occurrence is counted from the right to left. But `i=0` is not allowed though. Thus, to get the **second** occurrence of some pattern, use `i=2`, and to get the **second-last** occurrence, use `i=-2`.

The `stri_locate_ith(str, i, ...)` function uses the exact same argument and naming convention as `stringi`, to keep your code consistent. And just like `stringi::stri_locate_first/last`, the `stri_locate_ith(str, i, ...)` function is a vectorized function: `str` and  `i` as well as the pattern (`regex, fixed, coll, charclass`) can all be different-valued vectors. It is also vectorized in the sense that no loops (in R) are used, only vectorized functions.


&nbsp;

To transform the **second-last** occurrence, one can now use `stri_locate_ith()` in a very similar way as was done with `stri_locate_first/last`:

```{r}
x <- c("HELLO WORLD", "goodbye world")

loc <- stri_locate_ith( # this part is the key-difference
  x, -2, regex="a|e|i|o|u", case_insensitive=TRUE
)

extr <- stringi::stri_sub(x, from=loc)
repl <- chartr(extr, old = "a-zA-Z", new = "A-Za-z")
stringi::stri_sub_replace(x, loc, replacement=repl)
```

Notice that the code is virtually equivalent. We *only* need to change the locate function.

&nbsp;

## Substr - functions

The `tidyoperators` R-package includes the following "substr-" functions:

 - The `substr_repl(x, rp, ...)` function replaces a position (range) with string `rp`.
 - The `substr_chartr(x, old, new, ...)` function transforms the sub-string at a position (range) using `chartr(old, new)`. By default, it will translate upper-case characters to lower-case, and vice-versa.
 - The `substr_addin(x, addition, side, ...)` function adds the additional string `addition` at the side `side` of a position.
 - The `substr_extract(x, type, ...)` function extracts the string at, before, or after some position (range).
 - The `substr_arrange(x, arr, ...)` function re-arranges (sort, reverse, randomize) the sub-string at a position (range).

The "position" in the functions above can be specified either by giving the result of the `stri_locate_ith()` function (see the previous section) in argument `loc`, or one can give manual positions using the `start, end` or `at` arguments. 

Examples:

```{r}
x <- c("Good - good - GOOD", "Good - GOOD - good")
loc <- stri_locate_ith(
  # locate second "good" (ignore case) of each string in x:
  x, 2, regex="good", case_insensitive=TRUE 
)
```

```{r echo=FALSE}
tab <- data.frame(
  x=x,
  `substr_extract(x, loc=loc)` = substr_extract(x, loc=loc),
  `substr_extract(x, "before", loc=loc)` = substr_extract(x, "before", loc=loc),
  `substr_extract(x, "after", loc=loc)` = substr_extract(x, "after", loc=loc),
  `substr_repl(x, "??", loc=loc)` = substr_repl(x, "??", loc=loc),
  `substr_chartr(x, loc=loc)` = substr_chartr(x, loc=loc),
  `substr_addin(x, "~~", "after", loc=loc)` = substr_addin(x, "~~", "after", loc=loc),
  `substr_addin(x, "~~", "before", loc=loc)` = substr_addin(x, "~~", "before", loc=loc),
  `substr_arrange(x, loc=loc)` = substr_arrange(x, loc=loc),
  `substr_arrange(x, "decr", loc=loc)` = substr_arrange(x, "decr", loc=loc),
  `substr_arrange(x, "rev", loc=loc)` = substr_arrange(x, "rev", loc=loc),
  `substr_arrange(x, "rand", loc=loc)` = substr_arrange(x, "rand", loc=loc),
  check.names = FALSE
)
tab <- tab |> as.matrix() |> t()
kable(tab)
```


&nbsp;

# String infix operators

The `tidyoperators` R package implements infix operators for string arithmetic and sub-setting, as well some of their in-place modifier equivalents. For consistency, and to avoid masking other common operators, all string-related operators start with `%s`, where the "s" stands for "string".

&nbsp;


## String subsetting operators

As a first sub-setting operator, we have `x %sget% ss`, which returns a subset of each string in character vector `x`. Here `ss` is a vector of length 2, or a matrix with `nrow(ss)=length(x)` and 2 columns. The object `ss` should consist entirely of non-negative integers (thus 0, 1, 2, etc. are valid, but -1, -2, -3 etc are not valid). The first element/column of ss gives the number of characters counting from the left side to be extracted from x. The second element/column of ss gives the number of characters counting from the right side to be extracted from x.

Here are 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %sget% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %sget% ss
```

Thus `x %sget% ss` "gets" or extracts the given number of characters from the left and the right, and removes the rest.
There is also `x %strim% ss`, which is the opposite: it trims away the number of characters from the left and right as defined in the matrix `ss`, leaving you with whatever is left.

Here are again 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %strim% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %strim% ss
```


&nbsp; 


## String arithmetic

The `tidyoperators` package adds 4 arithmetic operators:

 - `x %s+% y` concatenates `x` and `y`;
 - `x %s-% p` removes pattern `p` from each string in character vector `x`;
 - `x %s*% n` repeats each string in character vector `x` for `n` times;
 - `x %s/% p` counts how often pattern `p` occurs in string or vector `x`
 
I.e.:

```{r}
"Hello "%s+% " world"
c("Hello world", "Goodbye world") %s-% " world"
c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 2:7
c("hello world & goodbye world", "world domination!") %s/% "world"
```

The right-side arguments `p`, `y`, and `n` can be a single value, or a vector of the same length as `x`.

&nbsp;

## Specifying Pattern search attributes in string infix operators

The `x %s-% p` and `x %s/% p` operators (and their in-place equivalents, given later), and the `%sgrep%` operator perform pattern matching for various purposes. By default the pattern matching is interpreted as case-sensitive `regex` patterns from `stringi`.

But, of course, sometimes one wants to change this. For example, one may want it to be case insensitive. Or perhaps one wants to use fixed expressions, or something else.

The `tidyoperators` package provides options for these cases. To use more refined pattern definition, simply replace the argument/right-hand-side expression `p` in the relevant operators with a call from the `s_pattern()` function.

The `s_pattern()` function uses the exact same argument convention as `stringi`. For example:

 - `s_pattern(regex=p, case_insensitive=FALSE, ...)`
 - `s_pattern(fixed=p, ...)`
 - `s_pattern(coll=p, ...)`
 - `s_pattern(charclass=p, ...)`

Note that the `s_pattern()` function only works for the infix operators (functions surrounded by `%` signs) related to strings in this R package.


Examples with Regular expressions:

```{r}
x <- c("Hello world", "Goodbye world")
p <- s_pattern(regex=" world")
x %s-% p

x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(regex="Ha")
x %s/% p
```

Examples with Fixed expressions:

```{r}
x <- c("Hello world", "Goodbye world")
p <- s_pattern(fixed=" world")
x %s-% p

x <- c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 10
p <- s_pattern(fixed="Ha")
x %s/% p
```

And so on. I'm sure you get the idea.

&nbsp;


# "Don't Repeat Yourself" - operators

## The transform_if function, and related operators

"Don't Repeat Yourself", sometimes abbreviated as "DRY", is the coding principle that you should try to reduce repeating patterns in your code (within reason).

Consider the following code:

`x <- ifelse(cond(x), f(x), g(x))`

Here a conditional subset of the object `x` is transformed where the condition is using a function referring to `x` itself. Consequently, reference to `x` is written **four times**!
If the object has a short name like "x, this doesn't matter too much. But if the object has a longer name like `very_long_name_1`, doing something like this:

```{r eval=FALSE}
very_long_name_1 <- ifelse(
  very_long_name_1>0,  log(very_long_name_1), very_long_name_1^2
)
```

becomes cumbersome quickly.

The tidyoperators package therefore adds the `transform_if(x, cond, trans_T, trans_F)` function which will tidy this up.
The above code can now be re-written as:

```{r eval=FALSE}
very_long_name_1 %<>% transform_if(\(x)x>0, log, \(x)^2)
```

Note that `x` must be a vector, matrix, or array.
Unlike `ifelse`, the transformations in `transform_if()` are evaluated as `trans_T(x[cond(x)])` and `trans_F(x[!cond(x)])`, ensuring no unnecessary warnings or errors occur.

Besides `transform_if`, the tidyoperators package also adds 2 "subset_if" operators:

 - The `x %[if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `TRUE`.

 - The `x %[!if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `FALSE`.

For example:

```{r}
object_with_very_long_name <- matrix(-10:9, ncol=2)
print(object_with_very_long_name)
object_with_very_long_name %[if]% \(x)x %in% 1:10
object_with_very_long_name %[!if]% \(x)x %in% 1:10
```


Another operator added by `tidyoperators` is `x %unreal =% y`, which replaces all NA, NaN, Inf and -Inf in `x` with the value given in `y`.

So `x %unreal =% y` is the same as `x[is.na(x)|is.nan(x)|is.infinite(x)] <- y`.


&nbsp;


## In-place modifying mathematical arithmetic


This R package includes infix operators for in-place modifying mathematical arithmetic.

Consider the following line of code:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

The same expression, `mtcars$mpg[mtcars$cyl>6]`, is written twice, making this code rather long and cumbersome, even though we're just squaring the expression. The well-known `magrittr` R-package has an in-place modifier pipe, `%<>%`. This works excellent for actual functions, like so:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% mean() # same as mtcars$mpg[mtcars$cyl>6] <- mean(mtcars$mpg[mtcars$cyl>6])
```
 
But for arithmetic, one needs to translate the operators into a function, and then perform the in-place modifier pipe:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %<>% raise_to_power(2)
```

This is better, but still not truly tidy.

This R package solves the above laid-out problem by implementing in-place modifying mathematical arithmetic for all mathematical operators, excluding matrix operators.

Here is a list of all in-place mathematical modifiers implemented in this R-package:

 - `x %+ =% y` is the same as`x <- x + y`;
 - `x %- =% y` is the same as`x <- x - y`;
 - `x %* =% y` is the same as`x <- x * y`;
 - `x %/ =% y` is the same as`x <- x / y`;
 - `x %^ =% p` is the same as`x <- x^p`;
 - `x %rt =% p` is the same as`x <- x^(1/p)`;
 - `x %logb =% b` is the same as`x <- log(x, base=b)`;
 - `x %alog =% b` is the same as`x <- b^x`; if `b=exp(1)`, this is the same as`x <- exp(x)`;
 - `x %alog =% exp(1)` is the same as exp(x)`.
 
All in-place modifying operators in this package end with " $\space$ =" (notice the extra space before the `=`). 

Lets look at the original problem:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

With `tidyoperators` one can now make this more tidy with the following:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %^ =% 2
```


&nbsp;


## In-place modifying string arithmetic and sub-setting

With the exception of  `%ss%`, all infix operators for string arithmetic and string sub-setting have their in-place modifying equivalent:

 - `x %s+ =% y` is the same as `x <- x %s+% y`
 - `x %s- =% p` is the same as `x <- x %s-% p`
 - `x %s* =% n` is the same as `x <- x %s*% n`
 - `x %s/ =% p` is the same as `x <- x %s/% p`
 - `x %sget =% ss` is the same as `x <- x %sget% ss`
 - `x %strim =% ss` is the same as `x <- x %strim% ss`


&nbsp;


# Import functions and operator

## import_as

One can load - but not attach - a package and assign it to an alias in base R using:

```{r eval=FALSE}
alias <- loadNamespace("packagename", lib.loc = lib.loc)
```

Doing the above, instead of attaching a package using `library()` or `require()`, can (sometimes) be quite beneficial: i.e. prevent overlapping namespaces, prevent overriding base/core R functions, prevent polluting the namespace, clarify which function came from which package.

Loading a package alias does have some drawbacks. One is that you cannot easily import multiple packages under the same alias. While one would probably not want to import **multiple** packages under a single alias most of the time, there may be a couple of situations where importing multiple packages into a single alias might be actually preferable:

 - A package may have one or several dependencies that are supposed to be loaded together. For example: the `tidytable` package essentially needs the `data.table` package to be loaded also.
 - A package may have many extensions you may want to load together. For example: `ggplot2` has many extensions (see https://exts.ggplot2.tidyverse.org/gallery/). If one wishes to alias `ggplot2` including some of its many extensions, one must be able to load multiple R packages under the same alias.
 - Suppose package `B` is supposed to overwrite a couple of functions from package `A` (for example if package `B` extends or improves the functionality from a function in package `A`). In that case you want to import package `A`, and then overwrite it with package `B`.

So there are several cases where it is perhaps desirable to load multiple packages under the same alias. And that is where `tidyoperator`'s `import_as()` function comes in. It allows loading multiple R packages under the same alias, and also informs the user which objects from a package will overwrite which objects from other packages, so you will never be surprised. The `import_as()` function also only loads exported functions (unlike `loadNamespace()`, which loads both internal and external functions). This is, I think, more desirable, as internal function should remain, you know, internal.

`import_as(alias, pkgs, lib.loc)` is thus essentially the multi-package equivalent of `alias <- loadNamespace(package, lib.loc)`.

Here is one example. Lets load  `data.table` and then `tidytable`, under the same alias, which I will call "tdt" (for "tidy data.table"):

```{r}
pkgs <- c("data.table", "tidytable")
import_as(fv, pkgs) # this creates the fv object
```


Now you can of course use those loaded packages as one would normally do when using a package alias.

A few notes on the usage of this operator when importing multiple R packages under the same alias:

 - The order of the character vector matters! If 2 packages share objects with the same name, the package named last will overwrite the earlier named package.
 - When supplying more than one package to the `import_as()` function, it is strongly advised to only import packages together under the same alias that are (reverse) dependencies of each other (i.e. they appear in each others Depends or Imports sections in the Description file).
 - Related to the previous point, the `import_as()` function only performs a very basic check for dependencies; the user is expected to use the `import_as()` function responsibly.

&nbsp;

## import_inops

When aliasing an R package, infix operators are also loaded in the alias. However, it may be cumbersome to use them from the alias. For example this:

```{r eval=FALSE}
to loadNamespace("tidyoperators")
to$`%row~%`(x, mat)
```


is very cumbersome. Therefore,  `tidyoperators` also adds the `import_inops(pkgs)` function, which exposes the infix operators from the packages specified in character vector `pkgs` to the current environment.

For example, the following code imports `data.table` and `tidytable`, and also exposes the infix operators from those packages to the (global) environment:

```{r}
pkgs <- c("data.table", "tidytable")
import_as(fv, pkgs) # this creates the fv object
import_inops(pkgs)
```

Just like in `import_as()`, the order of the packages matter whenever there's a conflict.

The `import_inops()` functions has the `exclude` and `include.only` arguments to specify exactly which infix operators to expose to the current environment. This can be handy to prevent overwriting any (user defined) infix operators already present in the current environment.

&nbsp;

## import_data

The `import_as()` function imports everything from the package namespace. But packages often also have data sets, which are often not part of the namespace.

The `data()` function in core R can already load data from packages, but this function loads the data into the global environment, instead of returning the data directly, making assigning the data to a specific variable a bit annoying.
Therefore, the `tidyoperators` package introduces the `import_data()` function, which directly returns a data set from a package.

For example, to import the `chicago` data set from the `gamair` R package, and assign it directly to a variable (without having to do re-assignment and so on), one simply runs the following:

```{r}
d <- import_data("chicago", "gamair")
head(d)
```

&nbsp;

## installed in - operator

The `pkgs %installed in% lib.loc` operator checks if one or more package(s) `pkgs` exist(s) in library location `lib.loc`.
Now you no longer have to attach a package with `require()` simply to check if it exists. Moreover, this operator makes it syntactically explicit in your code where you are looking for your R package(s).

Example:

```{r}
pkgs <- c("data.table", "tidytable")
pkgs %installed in% .libPaths()
```


Note that all "import_" functions in the `tidyoperators` package have a `lib.loc` argument to explicitly specify from where to get your packages (just like the `library()` function).

&nbsp;

# On libraries

## On date-based version control: the alternative to MRAN

As MRAN is no longer available, RStudio/Posit has taken up the mantle to provide date-based version control.

Simply go to https://packagemanager.rstudio.com/client/#/repos/2/overview. Then select a date and select your operating system (or, if you want to install from Source, leave it to Source). Scroll down to "Example repository setup code", and copy the code into R to use the repository.

For example, to set the repository URL in your current R script such that R packages from CRAN are downloaded and installed as they were available on 2 January 2023, when using a Microsoft Windows OS, simply run the following:

```{r}
options(repos = c(REPO_NAME = "https://packagemanager.rstudio.com/cran/2023-01-02"))
```

Using this in combination with project isolation (like using `force_libPaths()`) allows for complete control over your packages, ensuring your code will run even several years after you wrote your code.

&nbsp;

## force_libPaths (for simple Project Isolation)

The `renv` R package is the go-to R package for complex project isolation. But sometimes you just have a few simple R scripts, and only need quick and simple project isolation.

The base R's `.libPaths()` function sets the library paths where R looks for R packages when checking or loading/attaching R packages. As such, this function can be used for simple Project Isolation. In some occasions you may want to completely overwrite the libraries, including the system and site library paths. The `.libPaths()` functions does not allow the user to do that.

The `tidyoperators` package therefore adds the `force_libPaths()` function, as provided in https://milesmcbain.com/posts/hacking-r-library-paths/. The `force_libPaths()` function does allow forcing R to only use the library paths exactly as specified by the user.

Example:

```{r eval=FALSE}
force_libPaths("/mylibrary")
```

For more complex project isolation, please use `renv` instead. Do not use both `renv` and `force_libPaths()` (or even base R's `.libPaths()`) in the same project.


&nbsp;

# Speed and multi-threading

## stri_locate_ith

The `stri_locate_ith()` function needs to know the number of matches of some pattern for every string in a character vector, as well as the actual matches themselves, before being able to actually give the $i^\textrm{th}$ occurrence of some pattern. Thus `stri_locate_ith()` (at least in its current implementation) cannot be faster than the combined runtime of the `stri_locate_all()` and `stri_count()` functions. As `stri_locate_ith()` is written mostly only in fully vectorized statements in R (no loop), the function hardly takes more than twice the time of `stri_locate_all()` and `stri_count()` combined.

&nbsp;

## Substr-functions

All the string sub-setting functions have the `fish` argument, which is `FALSE` by default. If `fish=TRUE`, these functions will use `stringfish` functions instead of base R and `stringi` functions. The stringfish functions are usually faster, and also allow native multi-threading. Note that `stringfish` must be installed in order for this to work. And `stringfish` needs to be loaded also if you wish to also use multi-threading. Multi-threading in `stringfish` can be set-up by running `setoption(stringfish.nthreads=cl)` somewhere at the start of your code, where `cl` is the number of threads you want to use.

Don't use multi-threading unless you need to, as multi-threading has some overhead, thus its only faster with very large character vectors.

Example:

```{r}
x <- c("Goodmorning--Goodevening, and Goodnight",
       paste0(letters[1:13], collapse=""))
print(x)
loc <- stri_locate_ith(
  # locate second-last occurrence of "good" (ignore case) of each string in x:
  x, -2, regex="good", case_insensitive=TRUE 
)
substr_extract(x, loc=loc, fish = TRUE)
substr_extract(x, "before", loc=loc, fish = TRUE)
substr_extract(x, "after", loc=loc, fish = TRUE)
substr_repl(x, "??", loc=loc, fish = TRUE)
substr_chartr(x, loc=loc, fish = TRUE)
substr_addin(x, " ", "after", loc=loc, fish = TRUE) 
substr_addin(x, " ", "before", loc=loc, fish = TRUE) 
substr_arrange(x, loc=loc, fish = TRUE)
substr_arrange(x, "rev", loc=loc, fish = TRUE)
substr_arrange(x, "decr", loc=loc, fish = TRUE)
substr_arrange(x, "rand", loc=loc, fish = TRUE)

```


&nbsp;

# Recommended R packages

`stringi` is of course required for this packages. Besides that, I highly recommend using this R package alongside the 2 major operator-related R-packages, namely `magrittr` and `zeallot`.

For proper programming etiquette, I also highly recommend the following R packages:

 - The `this.path` R package (https://github.com/ArcadeAntics/this.path), which allows for getting the path of the current script (i.e. get the source file location), without all the issues found in packages such as the `here` and `whereami` packages.
 - The `rlang` R package, which one can use to get (among other things) more detailed back-traced errors, warnings, and messages.

&nbsp;

# Compatibility with other R packages

The `stringi` R package has the `%s+%` and `%s*%` operators. They do exactly the same things as in `tidyoperators`, and so the masking of these functions can safely be ignored. I also made sure not to name any of the operators in `tidyoperators` the same as the operators in `magrittr` and `zeallot`, so that should be safe also.

The `force_libPaths()` function is meant for super simple project isolation. One should probably not use the `force_libPaths()` function in a script when already using another system for project isolation like `renv` or `packrat`. In fact, you probably shouldn't use even base R's `.libPaths()` function when also using `renv`. It's either one or the other. The rest of the functions and operators in `tidyoperators` are of course fully compatible with `renv`.

Note that, by default, `renv` only registers packages loads using plain `library()` or `require()` calls. Anything different from that, even things like `for(... in ...)library(...)` or `if(...)library(...)`, will not be understood by `renv`. Therefore, if using `renv`, please make sure to set the following:

```{r eval=FALSE}
renv::settings$snapshot.type("all")
```

This will make sure that all packages installed in your project library, regardless of how they are loaded,  will all be registered by `renv`. This makes `renv` compatible with calls like  `import_as` from `tidyoperators`, and things like `for(... in ...)library(...)` or `if(...)library(...)`.

&nbsp;

# Conclusion

I hope this R package will make your life a little bit tidier.

