---
title: "Safer decimal (in)equality testing"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Safer decimal (in)equality testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinyoperations)
```


This package adds the `%d==%, %d!=% %d<%, %d>%, %d<=%, %d>=%` (in)equality operators, which perform safer decimal number truth testing. They are virtually equivalent to the regular (in)equality operators,
`==, !=, <, >, <=, >=`,
except for one aspect: The decimal number truth testing operators assume that if the absolute difference between `x` and `y` is smaller than the Machine tolerance, `sqrt(.Machine$double.eps)`, then `x` and `y` ought to be consider to be equal.

Thus these provide safer decimal number (in)equality operators.

For example: `(0.1*7) == 0.7` returns `FALSE`, even though they are equal, due to the way decimal numbers are stored in programming languages like `R`, `Python`, etc. But `(0.1*7) %d==% 0.7` returns `TRUE`.

Some examples:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)
x == y # gives FALSE, but should be TRUE
x!= y # gives TRUE, should be FALSE
x > y # not wrong
x < y # gives TRUE, should be FALSE
x %d==% y # here it's done correctly
x %d!=% y
x %d<% y # correct
x %d>% y # correct
x %d<=% y # correct
x %d>=% y # correct

```

&nbsp;

There are also the `x %d{}% bnd` and `x %d!{}% bnd` operators, where `bnd` is a vector of length 2, or a 2-column matrix `(nrow(bnd)==length(x) or nrow(bnd)==1)`.
The `x %d{}% bnd` operator checks if `x` is within the **closed** interval with bounds defined by `bnd`.
The `x %d!{}% bnd` operator checks if `x` is outside the **closed** interval with bounds defined by `bnd`.


Examples:

```{r}

x <- c(0.3, 0.6, 0.7)
bnd <- matrix(c(x-0.1, x+0.1), ncol=2)
x %d{}% bnd
x %d!{}% bnd

```

&nbsp;

