---
title: "Safer functionality"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Safer functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
```


# Decimal (in)equality testing operators

This package adds the `%d==%, %d!=% %d<%, %d>%, %d<=%, %d>=%` (in)equality operators, which perform safer decimal number truth testing. They are virtually equivalent to the regular (in)equality operators,
`==, !=, <, >, <=, >=`,
except for 2 aspects:

 1) The `%d...%` operators assume that if the absolute difference between any two numbers x and y is smaller than the Machine tolerance, sqrt(.Machine$double.eps), then x and y should be consider to be equal.
 For example: `0.1*7 == 0.7` returns `FALSE`, even though they are equal, due to the way decimal numbers are stored in programming languages like 'R' and 'Python'. But `0.1*7 %d==% 0.7` returns `TRUE`.

 2) Only numeric input is allowed, so characters are not coerced to numbers.
 I.e. `1 < "a"` gives `TRUE` , whereas `1 %d<% "a"` gives an error.
 For character equality testing, see %s==% from the 'stringi' package.


Thus these provide safer decimal number (in)equality operators.



Some examples:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)
x == y # gives FALSE, but should be TRUE
x!= y # gives TRUE, should be FALSE
x > y # not wrong
x < y # gives TRUE, should be FALSE
x %d==% y # here it's done correctly
x %d!=% y
x %d<% y # correct
x %d>% y # correct
x %d<=% y # correct
x %d>=% y # correct

```

&nbsp;

There are also the `x %d{}% bnd` and `x %d!{}% bnd` operators, where `bnd` is a vector of length 2, or a 2-column matrix `(nrow(bnd)==length(x) or nrow(bnd)==1)`.
The `x %d{}% bnd` operator checks if `x` is within the **closed** interval with bounds defined by `bnd`.
The `x %d!{}% bnd` operator checks if `x` is outside the **closed** interval with bounds defined by `bnd`.


Examples:

```{r}

x <- c(0.3, 0.6, 0.7)
bnd <- cbind(x-0.1, x+0.1)
x %d{}% bnd
x %d!{}% bnd

```


&nbsp;

# Safer atomic type casting

Atomic type casting in R is generally performed using the functions `as.logical()`, `as.integer()`, `as.double()`, `as.character()`.

These functions have the annoying property that they strip attributes. I've taken the liberty of creating more convenient atomic conversion functions:

 - `as_bool()`: same as `as.logical()`, but with attributes preserved.
 - `as_int()`: same as `as.integer()`, but with attributes preserved.
 - `as_dbl()`: same as `as.double()` (i.e. convert to real numbers), but with attributes preserved.
 - `as_chr()`: same as `as.character()`, but with attributes preserved.
 
All attributes except the "class" attribute are preserved.

Examples:

```{r}
x <- c(rep(0, 2), seq(0, 2.5, by=0.5)) |> matrix(ncol=2)
colnames(x) <- c("one", "two")
attr(x, "test") <- "test"
print(x)

as_bool(x)
as_int(x)
as_dbl(x)
as_chr(x)
```



&nbsp;


# Locked constants

One can re-assign the values `T` and `F`. One can even run something like `T <- FALSE` and `F <- TRUE`! `tinycodet` adds the `lock_TF()` function that forces `T` to stay `TRUE` and `F` to stay `FALSE`. Essentially, the `lock_TF()` function creates the locked constant `T` and `F`, assigned to `TRUE` and `FALSE` respectively, to prevent the user from re-assigning them. Removing the created `T` and `F` constants allows re-assignment again.

&nbsp;

The `X %<-c% A` operator creates a `constant` `X` with assignment `A`. Constants cannot be changed, only accessed or removed. So if you have a piece of code that requires some unchangeable constant, use this operator to create said constant.


&nbsp;


# form()
## (new in version 0.4.5; under development)

When creating a formula with the tilde (` ⁠~` ⁠) operator, and storing a formula in a variable to be used later, the environment is captured by the formula.
Therefore, any object in the captured environment might not be freed from the memory, potentially creating some memory leak.


The `form()` function is a convenience function, to quickly create and return/store a formula more safely, by having no default environment, and allowing the user to specify the environment explicitly.

It can also quickly convert a single string to a formula, again allowing for explicit environment specification, and having no environment by default.

```{r}
# see also http://adv-r.had.co.nz/memory.html


f1 <- function() {
  foo <- c(letters, LETTERS)
  return(10)
}
x1 <- f1()
environment(x1) |> as.list() # empty, since no formula is used inside f1(), so safe


f2 <- function() {
  foo <- c(letters, LETTERS)
  out <- a ~ b
  return(out)
}
x2 <- f2()
environment(x2) |> as.list() # NOT safe: contains all objects from f2()
exists("foo", envir = environment(x2)) # = TRUE: "foo" still exists!
environment(x2)$foo # can still access it; probably won't be removed by gc()


f3 <- function() {
  foo <- c(letters, LETTERS)
  return(form(a ~ b))
}
x3 <- f3()
environment(x3) |> as.list() # empty, since form() is used, so safe


f4 <- function() {
  foo <- c(letters, LETTERS)
  return(form("a ~ b"))
}
x4 <- f4()
environment(x4) |> as.list() # empty, since form() is used, so safe

```


&nbsp;

# with_pro and aes_pro
## (new in version 0.4.5; under development)

'tinycodet' provides standard-evaluated versions of the common quoting functions `with()` and `ggplot2::aes()`: `with_pro()` and `aes_pro()`, respectively.

```{r fig.width = 8}

requireNamespace("ggplot2")


d <- import_data("ggplot2", "mpg")

# mutate data:
myform <- form(~ displ + cyl + cty + hwy)
d$mysum <- with_pro(d, myform)
summary(d)

# plotting data:
x <- form("cty")
y <- form(~ sqrt(hwy))
color <- form(~ drv)

ggplot2::ggplot(d, aes_pro(x, y, color = color)) +
  ggplot2::geom_point()


```

&nbsp;

# Safer Partial Matching
## (new in version 0.4.5; under development)

The `safer_partialmatch()`  forces 'R' to give a warning when partial matching occurs when using the dollar ($) operator, or when other forms of partial matching occurs. It simply calls the following:

```{r echo = TRUE, eval = FALSE}

options(
   warnPartialMatchDollar = TRUE,
   warnPartialMatchArgs = TRUE,
   warnPartialMatchAttr = TRUE
 )

```


&nbsp;

