---
title: "Import system - main functions"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Import system - main functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
```

&nbsp;

# Introduction

One can load a package without attaching the package (i.e. using `::` or using a package alias), or one can attach a package (i.e. using `library()` or `require()`).

The advantages and disadvantages of loading without attaching a package versus attaching a package - at least those relevant for this article - can be compactly presented in the following table:

```{r echo=FALSE, results='markup'}
X <- rbind(
  # c(aspect, alias, attach)
  c("prevent masking functions from other packages", "Yes (+)", "No (-)"),
  c("prevent masking core R functions", "Yes (+)", "No (-)"),
  c("clarify which function came from which package", "Yes (+)", "No (-)"),
  c("place/expose functions only in current environment instead of globally", "Yes (+)", "No (-)"),
  c("prevent namespace pollution", "Yes (+)", "No (-)"),
  c(
  "minimize typing - especially for infix operators <br> (i.e. typing ``package::`%op%`(x, y)`` instead of `x %op% y` is cumbersome)", "No (-)", "Yes (+)"),
  c("use multiple related packages, <br> without constantly switching between package prefixes", "No (-)", "Yes (+)"),
  c("NOTE: + = advantage, - = disadvantage", "", "")
)
colnames(X) <- c("aspect", "alias / ::", "attaching")
X <- as.data.frame(X)
rownames(X) <- c(1:7, "")
knitr::kable(X, "html", row.names = TRUE, escape = FALSE)
```


What `tinycodet` attempts to do with its import system, is to somewhat find the best of both worlds. Basically, `tinycodet` has functions that allow the following import functionality lacking in base R:

 - Allow an R package + its re-exports + its dependencies + its extensions to be loaded under **one alias**. This essentially combines the attaching advantage of using multiple related packages (row 7 in table above), whilst keeping most advantages of aliasing a package.
 - Allow **exposing infix operators** to the **current environment**. This gains the attaching advantage of less typing (row 6 in table above), whilst simultaneously avoiding the disadvantage of attaching functions from a package globally (row 4).

The import package system presented here is just another option provided, just like the [import](https://github.com/rticulate/import) and [box](https://github.com/klmr/box) packages provide their own alternative import systems. Please feel free to completely ignore this article if you're really adamant on attaching packages using `library()`/`require()` :-).

This article is rather lengthy, so I will start with a quick example code using `tinycodet`' import system:

```{r}
# loading "tidytable" + its re-exports + "data.table" under alias "tdt.":
import_as( 
  ~ tdt., "tidytable", dependencies = "data.table"
)

# exposing operators from `magrrittr` to current environment:
import_inops("magrittr")

# directly assigning the "starwars" dataset to object "d":
d <- import_data("dplyr", "starwars") 

# see it in action:
d %>% tdt.$filter(species == "Droid") %>%
  tdt.$select(name, tdt.$ends_with("color"))

rm(list=ls()) # clearing everything
```

The above code is run *without attaching* any of the packages or its dependencies. So none of the problems with attaching a package is present.

Despite the length of this article, which is mostly due to me being overly detailed, the import system is made to be *very simple for the user*.

What follows are descriptions of the main functions that together form this new, infix-operator friendly **&** multi-package assignment friendly, import management system.

&nbsp;

# import_as

One can load a package without attaching it, and assign it to an alias, in base R, using, for example:

```{r eval=FALSE}
alias <- loadNamespace("packagename", lib.loc = lib.loc)
```

Doing the above, instead of attaching a package using `library()` or `require()`, can (often) be quite beneficial (see Introduction section in this article).

Loading a package in an alias, whether one does it through base R or through an R package like `import` or `box`, does have some drawbacks. One is that you cannot easily import multiple related packages under the same alias. Although one may not *always* want to import *multiple* related packages under a *single alias*, there are several situations where importing multiple packages into a single alias may be preferable:

 - If multiple packages are meant to work together, constantly switching between the different package name/alias prefixes may eventually become annoying and even syntactically chaotic-looking.
 - A package may have one or several dependencies that are supposed to be loaded together. For example: the `tidytable` package essentially needs the `data.table` package to be loaded also.
 - A package may have many extensions you may want to load together. For example: `ggplot2` has many extensions (see https://exts.ggplot2.tidyverse.org/gallery/). If one wishes to alias `ggplot2` including some of its many extensions, it would be nice to load multiple R packages under the same alias.
 - Suppose package `B` is supposed to overwrite a couple of functions from package `A` (for example if package `B` extends or improves the functionality from a function in package `A`). In that case you may want to import package `A`, and then overwrite it with package `B`.

So there are several cases where it is perhaps desirable to load *multiple* related packages under one alias.

And that is where `tinycodet`'s `import_as()` function comes in. It loads an R package + its re-exports under an alias, and also loads any specified **dependencies** and/or **extensions** of the package under the very same alias. It also informs the user which objects from a package will overwrite which objects from other packages, so you will never be surprised.

`import_as(alias, ...)` is thus essentially the more user-friendly and multi-package equivalent of `alias <- loadNamespace(...)`.

The main arguments of the `import_as()` function are:

 - `alias`: the name of the alias object under which to load the package(s). Can be given as a single string or as a formula with a single term. To keep aliases easily distinguishable from other objects that can also be subset with the `$` operator, I recommend ending (not starting!) all alias names with a dot (.).
 - `main_package`: the name (string) of the main package to load.
 - `re_exports`: Some R packages export functions that are not defined in their own package, but in their direct dependencies - "re-exports". If `TRUE` (default), the re-exports of the `main_package` are added to the alias, analogous to the behaviour of base R's `::` operator. If `FALSE`, re-exports are not added.
 - `dependencies`: an optional character vector giving the dependencies of the `main_package` to load under the alias also.
 - `extensions`: an optional character vector giving the extensions/reverse-dependencies of the `main_package` to load under the same alias also.
 - `lib.loc`: the library paths to look for the packages; defaults to `.libPaths()`. This argument is present in all `import_` - functions.

Here is one example. Lets load  [data.table](https://github.com/Rdatatable/data.table) and its extensions [tidytable](https://github.com/markfairbanks/tidytable), under the same alias, which I will call "tdt." (for "tidy data.table"):

```{r}
import_as(~ tdt., "data.table", extensions = "tidytable") # this creates the tdt. object
```


Now you can of course use those loaded packages as one would normally do when using a package alias: `tdt.$some_function()`.

&nbsp;

# import_inops

When aliasing an R package, infix operators are also loaded in the alias. However, it may be cumbersome to use them from the alias. For example this:

```{r eval=FALSE}
import_as(~ to., "tinycodet")
to.$`%row~%`(x, mat)
```

or this:

```{r eval=FALSE}
tinycodet::`%row~%`(x, mat)
```


is very cumbersome.

Therefore, `tinycodet` also adds the `import_inops()` function, which exposes the infix operators. The infix operators are exposed to the current environment (like the global environment, or the environment within a function), but does not attach the functions to the namespace.

For example, to expose the infix operators in the alias object from before to the current environment, one can do the following:

```{r}
import_inops(expose = tdt.)
```

One can give the `unexpose` argument instead of the `expose` argument, which will delete the infix operators from those packages/package alias exposed in the current environment by `import_inops()`. Infix operators defined by the user will not be touched. For example:

```{r}
import_inops(unexpose = tdt.)
```


One can also expose and unexpose the infix operators directly from a package, instead of via an alias object. In that case the package name must be given as a string.

For example, the following code exposes the infix operators from the [data.table](https://github.com/Rdatatable/data.table) R package:

```{r}
import_inops(expose ="data.table")
```

And similarly one can remove the exposed infix operators again from the current environment as follows:

```{r}
import_inops(unexpose = "data.table")
```


&nbsp;

The `import_inops()` function has the `exclude` and `include.only` arguments to specify exactly which infix operators to expose to the current environment, as well as the `overwrite` and `inherits` arguments to specify what to do when the infix operators you are about to expose already exist in the current environment (and loaded namespaces). This can be handy to prevent overwriting any (user defined) infix operators already present in the current environment or loaded namespaces.

Examples:

```{r}
import_inops(expose = tdt., include.only = ":=")
import_inops(unexpose = tdt.)
import_inops(expose = "data.table", , include.only = ":=")
import_inops(unexpose = "data.table")
```


&nbsp;

If the user would rather attach the infix operators to the (global) namespace, `tinycodet` provides the `pkg_lsf()` function, which returns a character vector listing all functions or infix operators from a package. This vector can then be used in the `include.only` argument of the `library()` function. Like so:

```{r eval=FALSE}
library(magrittr, include.only = pkg_lsf("magrittr", type = "inops"))
```


&nbsp;

# import_data

The `import_as()` and `import_inops()` functions get all functions from the package namespace. But packages often also have data sets, which are often not part of the namespace.

The `data()` function in core R can already load data from packages, but this function loads the data into the global environment, instead of returning the data directly, making assigning the data to a specific variable a bit annoying.
Therefore, the `tinycodet` package introduces the `import_data()` function, which directly returns a data set from a package.

For example, to import the `chicago` data set from the [gamair](https://github.com/cran/gamair) R package, and assign it directly to a variable (without having to do re-assignment and so on), one simply runs the following:

```{r}
d <- import_data("gamair", "chicago")
head(d)
```


&nbsp;

# Example

One R package that could benefit from the import system introduced by `tinycodet`, is the [dplyr](https://github.com/tidyverse/dplyr) R package. The [dplyr](https://github.com/tidyverse/dplyr) R package overwrites **core R** functions (including base R) and it overwrites functions from pre-installed recommended R packages (such as `MASS`). I.e.:

```{r}
rm(list=ls()) # clearing environment again
library(MASS)
library(dplyr) # <- notice dplyr overwrites base R and recommended R packages

# detaching dplyr again:
detach("package:dplyr")
```

Moreover, [dplyr](https://github.com/tidyverse/dplyr)'s function names are sometimes generic enough that there is no obvious way to tell if a function came from [dplyr](https://github.com/tidyverse/dplyr) or some other package (for comparison: one can generally recognize `stringi` functions as they all start with `stri_`). If you look at the CRAN page for [dplyr](https://github.com/tidyverse/dplyr), you'll notice it has some interesting extensions you might want to use, such as [powerjoin](https://github.com/moodymudskipper/powerjoin).

To prevent masking base R functions, and to prevent obscurity regarding which functions come from [dplyr](https://github.com/tidyverse/dplyr) and [powerjoin](https://github.com/moodymudskipper/powerjoin), and which functions come from core R, one could constantly use `dplyr::` and `powerjoin::`. But constantly switching between package prefixes or aliases is perhaps undesirable. 

So here `tinycodet`' `import_as()` function might help. Below is an example where [dplyr](https://github.com/tidyverse/dplyr) is loaded (including its re-exports), along with [powerjoin](https://github.com/moodymudskipper/powerjoin) (which is an extension), all under one alias which I'll call "`dpr.`". Moreover, all infix operators from `magrittr` are exposed to the current environment.

```{r}
import_as(
  ~ dpr., "dplyr", extensions = "powerjoin", lib.loc = .libPaths()
)

import_inops("magrittr") # getting the infix operators from `magrittr`
```

The functions from [dplyr](https://github.com/tidyverse/dplyr) can now be used with the `dpr.$` prefix. This way, base R functions are no longer overwritten, and it will be clear for someone who reads your code whether functions like the `filter()` function is the base R filter function, or the [dplyr](https://github.com/tidyverse/dplyr) filter function, as the latter would be called as `dpr.$filter()`.

Let's first run a simple example code with the loaded functions:

```{r}
d <- import_data("dplyr", "starwars")
d %>%
  dpr.$filter(.data$species == "Droid") %>% # notice the ".data" pronoun can be used without problems
  dpr.$select(name, dpr.$ends_with("color"))
```

Just add `dpr.$` in front of the functions you'd normally use, and everything works just as expected.

Now lets run an example from the [powerjoin](https://github.com/moodymudskipper/powerjoin) GitHub page (https://github.com/moodymudskipper/powerjoin), using the above alias:

```{r}
male_penguins <- dpr.$tribble(
     ~name,    ~species,     ~island, ~flipper_length_mm, ~body_mass_g,
 "Giordan",    "Gentoo",    "Biscoe",               222L,        5250L,
  "Lynden",    "Adelie", "Torgersen",               190L,        3900L,
  "Reiner",    "Adelie",     "Dream",               185L,        3650L
)

female_penguins <- dpr.$tribble(
     ~name,    ~species,  ~island, ~flipper_length_mm, ~body_mass_g,
  "Alonda",    "Gentoo", "Biscoe",               211,        4500L,
     "Ola",    "Adelie",  "Dream",               190,        3600L,
"Mishayla",    "Gentoo", "Biscoe",               215,        4750L,
)
dpr.$check_specs()

dpr.$power_inner_join(
  male_penguins[c("species", "island")],
  female_penguins[c("species", "island")]
)

```


Notice that the only change made, is that all functions start with `dpr.$`, the rest is the same. No need for constantly switching between `dplyr::...`, `powerjoin::...` and so on - yet it is still clear from the code that the functions came from the [dplyr](https://github.com/tidyverse/dplyr) + [powerjoin](https://github.com/moodymudskipper/powerjoin) family, and there is no fear of overwriting functions from other R packages - let alone core R functions.

&nbsp;
