---
title: "Import system - additional details"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Import system - additional details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
```

&nbsp;

# Introduction

The previous article, "Import system - main functions", discussed the main functions of the import system. Please read that article first before reading this article.

&nbsp;

&nbsp;

# S3/S4 methods: they just work

When importing packages with `tinycodet`' import system, S3 and S4 methods will work just fine.

For example, the following code with S3 just works:

```{r}
import_as(.dpr ~ dplyr)
import_from("magrittr", import_ls("magrittr", "infix"))
d <- import_data("dplyr", "starwars")
d <- d %>% .dpr$group_by(species)

isS3method(f="arrange", class="data.frame", envir = .dpr) # this is an S3 method
isS3method(f="relocate", class="data.frame", envir = .dpr) # this is an S3 method
# this works:
d %>%
  .dpr$arrange(.dpr$desc(mass)) %>%
  .dpr$relocate(species, mass)
```

and so do these S4 methods:

```{r}
import_from("broadcast", "bc.d")

bc.d(1:10, array(1:10, c(1,10)), "+") # fast broadcasted addition

```


So when importing packages, everything works as expected, including S3 and S4 methods.


&nbsp;

# Alias attributes

The `attr.import()` function allows the user to access the special attributes stored and locked inside the alias object. These attributes show which imported package overwrites which imported functions, in what order the packages are imported and so on.

Here are some examples.

Show the packages imported under the alias, and in which order the packages are imported, and from which packages the re-exported functions came from:

```{r}
import_as(.tdt ~ tidytable, deps = "data.table")

attr.import(.tdt, "pkgs")
```

Show which functions from which packages "win" conflicts:

```{r}
attr.import(.tdt, "conflicts")|> knitr::kable()
```


The help file on `attr.import()` provides more details on each of these options.

&nbsp;


# help.import

The `help.import()` function gets the help file for a function `i` (or topic string `i`), even if the function is inside an alias object, or if the function is an unattached function (like exposed infix operators).

Example:

```{r eval=FALSE}
import_as(.dpr ~ "dplyr")
import_from("magrittr", import_ls("magrittr", "infix"))

help.import(i = .mr$add)
help.import(i = `%>%`)
help.import(i = "add", alias = .mr)
```


&nbsp;

# Miscellaneous comments on package imports


The [magrittr](https://github.com/tidyverse/magrittr) and [rlang](https://github.com/r-lib/rlang) packages add "pronouns" to R: `.`, `. data`, `.env`. Fret not, for pronouns work regardless if you attached a package or not. And you don't need to use something like `rlang::.data` or `rlang.$.data` for a pronoun to work. They just work.

&nbsp;


There are some additional miscellaneous functions related to the package import system that should perhaps be mentioned also:

 - the `pkg_get_deps()` function gets the dependencies (or the enhances) of a package, regardless if the package is CRAN or non-CRAN. See the help file for details.
 - the `pkgs %installed in% lib.loc` operator checks if the packages specified in character vector `pkgs` are installed in library paths `lib.loc`, and does this without attaching or even loading the packages.


&nbsp;


