---
title: "String related infix operators"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{String related infix operators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
```


&nbsp;

# Overview

'tinycodet' adds 3 sets of string-related operators.

First, sub-setting operators:

 - `x %sget% ss`: returns the first `n1` and last `n2` characters from each string in character vector `x`.
 - `x %strim% ss` trims away first `n1` and last `n2` characters from each string in character vector `x`.


Second, 'stringi' already has the `%s+%`, `%s*%`, and `%s$%` operators, and 'tinycodet' adds some additional string arithmetic operators to complete the set:

 - `x %s-% p` removes pattern `p` from each string in character vector `x`;
 - `x %s/% p` counts how often pattern `p` occurs in each string of character vector `x`.
 - `x %s//% brk` counts how often the text boundary specified in list `brk` occurs in each string of character vector `x`.
 - `x %ss% p` splits the strings in `x` by a delimiter character/pattern defined in `p`, and removes `p` in the process.


And finally, string search operators:

 - `x %s{}% p` operator checks for every string in character vector `x` if the pattern defined in `p` is present. Can also be used to check if the strings specifically start or end with pattern `p`.
 - `x %s!{}% p` operator checks for every string in character vector `x` if the pattern defined in `p` is **not** present. Can also be used to check if the strings specifically **does not** start or end with pattern `p`.
 - `strfind()<-` locates, extracts, or replaces found patterns.
 
&nbsp;

The `x %s-% p` and `x %s/% p` operators, and the string detection operators (`%s{}%`, `%s!{}%`, `strfind()<-`) perform pattern matching for various purposes. When a character vector or string is given on the right hand side, this is interpreted as case-sensitive `regex` patterns from `stringi`.

But, of course, sometimes one wants to change this. For example, one may want it to be case insensitive. Or perhaps one wants to use fixed expressions, or something else.

Instead of giving a string or character vector of regex patterns, one can also supply a list to the right-hand side, to specify exactly how the pattern should be interpreted. The list should use the exact same naming convention as `stringi`. For example:

 - `list(regex=p, case_insensitive=FALSE, ...)`
 - `list(fixed=p, ...)`
 - `list(coll=p, ...)`
 - `list(charclass=p, ...)`

For convenience, 'tinycodet' adds the following functions for this purpose:

 - `s_regex(p, ...)` is equivalent to `list(regex = p, ...)`
 - `s_fixed(p, ...)` is equivalent to `list(fixed = p, ...)`
 - `s_coll(p, ...)` is equivalent to `list(coll = p, ...)`
 - `s_chrcls(p, ...)` is equivalent to `list(charclass = p, ... )`


The next sections will give more details on the given overview.


&nbsp;

&nbsp;


# String subsetting operators

The `x %sget% ss` operator returns a subset of each string in character vector `x`. Here `ss` is a vector of length 2, or a matrix with `nrow(ss)=length(x)` and 2 columns. The object `ss` should consist entirely of non-negative integers (thus 0, 1, 2, etc. are valid, but -1, -2, -3 etc are not valid). The first element/column of ss gives the number of characters counting from the left side to be extracted from x. The second element/column of ss gives the number of characters counting from the right side to be extracted from x.

Here are 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %sget% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %sget% ss
```

Thus `x %sget% ss` "gets" or extracts the given number of characters from the left and the right, and removes the rest.
There is also `x %strim% ss`, which is the opposite: it trims away the number of characters from the left and right as defined in the matrix `ss`, leaving you with whatever is left.

Here are again 2 examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(2,3)
x %strim% ss

x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
ss <- c(1,0)
x %strim% ss
```



&nbsp;


# String arithmetic

The `tinycodet` package includes 7 string arithmetic operators (3 of which re-exported from 'stringi'):

 - `x %s+% y` concatenates `x` and `y` (exported from 'stringi');
 - `x %s-% p` removes pattern `p` from each string in character vector `x`;
 - `x %s*% n` repeats each string in character vector `x` for `n` times (exported from 'stringi');
 - `x %s/% p` counts how often pattern `p` occurs in each string of character vector `x`.
 - `x %s//% brk` counts how often the text boundary specified in list `brk` occurs in each string of character vector `x`.
 - `e1 %s$% e2` provides access to `stri_sprintf` (exported from 'stringi');
 - `x %ss% p` splits the strings in `x` by a delimiter character/pattern defined in `p`, and removes `p` in the process.
 
I.e.:

```{r}
"Hello "%s+% " world"
c("Hello world", "Goodbye world") %s-% " world"
c("Hello world", "Goodbye world") %s-% s_fixed(" world")
c("Ha", "Ho", "Hi", "Hu", "He", "Ha") %s*% 2:7
c("hello World & goodbye world", "world domination!") %s/% s_fixed("world", case_insensitive = TRUE)
c("hello world & goodbye world", "world domination!") %s//% list(type = "word")
```

The right-side arguments `y`, and `n` can be a single value, or a vector of the same length as `x`.
The right-side argument `p` can be string or character vector, or a list as described in the Overview section.

&nbsp;

# String search operators

The `x %s{}% p` operator checks for every string in character vector `x` if the pattern defined in `p` is present.
The `x %s!{}% p` operator checks for every string in character vector `x` if the pattern defined in `p` is NOT present.

Examples:

```{r}
x <- c(paste0(letters[1:13], collapse=""), paste0(letters[14:26], collapse=""))
print(x)
x %s{}% "a"
x %s!{}% "a"
which(x %s{}% "a")
which(x %s!{}% "a")
x[x %s{}% "a"]
x[x %s!{}% "a"]
```


When supplying a list on the right hand side (see the Overview section above), one can include the list element `at = "start"` or `at = "end"`:

 * Supplying `at = "start"` will check if the pattern appears at the start of a string (see `stringi::stri_startswith`).
 * Supplying `at = "end"` will check if the pattern appears at the end of a string (see `stringi::stri_endswith`).

Examples:

```{r}

x <- c(paste0(letters, collapse=""), paste0(rev(letters), collapse=""), NA)
p <- s_fixed("abc", at = "start")
x %s{}% p
stringi::stri_startswith(x, fixed = "abc") # same as above

p <- s_fixed("xyz", at = "end")
x %s{}% p
stringi::stri_endswith(x, fixed = "xyz") # same as above

p <- s_fixed("cba", at = "end")
x %s{}% p
stringi::stri_endswith(x, fixed = "cba") # same as above

p <- s_fixed("zyx", at = "start")
x %s{}% p
stringi::stri_startswith(x, fixed = "zyx") # same as above

```


`strfind()<-` locates, extracts, or replaces found patterns. Like the other operators, the argument `p` can be a string or character vector, or a list as described in the Overview section above.

Examples:

```{r}

x <- rep('The quick brown fox jumped over the lazy dog.', 3)
print(x)
p <- s_fixed(c('quick', 'brown', 'fox'))
rp <- c('SLOW',  'BLACK', 'BEAR')
x %s{}% p
strfind(x, p)
strfind(x, p) <- rp
print(x)

```




&nbsp;

