---
title: "Don't Repeat Yourself"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Don't Repeat Yourself}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinyoperations)
```


# The transform_if function

"Don't Repeat Yourself", sometimes abbreviated as "DRY", is the coding principle that you should try to reduce repeating patterns in your code (within reason).

Consider the following code:


```{r echo=TRUE}
object <- matrix(c(-9:8, NA, NA) , ncol=2)
y <- 0
z <- 1000
ifelse(
  is.na(object>y), -z,
  ifelse(
    object>y,  log(object), object^2
  )
)
```


Here a conditional subset of the object `object` is transformed where the condition is using a function referring to `object` itself. Consequently, reference to `object` is written 4 times! This can become cumbersome quickly. Notice also that the above code gives an unnecessary warning, due to `ifelse()` requiring the entirety of `log(object)`.


The `tinyoperations` package therefore adds the `transform_if(x, cond, yes, no, other)` function, which will "dry" this up.
Here, in argument `cond` a function must be given that returns a logical vector. For every value where `cond(x)==TRUE`, function `yes(x)` is run, for every value where `cond(x)==FALSE`, function `no(x)` is run, and for every value where `cond(x)==NA`, function `other` is run. Because a function-based approach is used instead of directly supplying vectors, unnecessary warnings and annoying errors are avoided (unlike the above code).

The above code can now be re-written in a less warning/error prone and more compact manner as:

```{r echo=TRUE}
object |> transform_if(\(x)x>y, log, \(x)x^2, \(x) -z)
```


Instead of supplying a function for `cond`, one can also directly supply a logical vector to argument `cond`. Moreover, when the transformed value is an atomic scalar, you don't really need a function; you can just fill in the scalar (vectors are not allowed though, as that will lead the same unnecessary warnings or even annoying errors as occur with `ifelse()`).

So one can thus also re-write the original code (without warnings/errors and more compact) as:

```{r echo=TRUE}
object |> transform_if(object > y, log, \(x)x^2, -z)
```

&nbsp;

# Subset if and unreal replacement

The `tinyoperations` package adds 2 "subset_if" operators:

 - The `x %[if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `TRUE`.

 - The `x %[!if]% cond` operator selects elements from vector/matrix/array `x`, for which the result of `cond(x)` returns `FALSE`.

For example:

```{r}
object_with_very_long_name <- matrix(-10:9, ncol=2)
print(object_with_very_long_name)
object_with_very_long_name %[if]% \(x)x %in% 1:10
object_with_very_long_name %[!if]% \(x)x %in% 1:10
```

&nbsp;

Another operator added by `tinyoperations` is `x %unreal =% y`, which replaces all NA, NaN, Inf and -Inf in `x` with the value given in `y`.

So `x %unreal =% y` is the same as `x[is.na(x)|is.nan(x)|is.infinite(x)] <- y`.


&nbsp;


# Generalized in-place modifier


This R package includes a general in-place modifying infix operator.

Consider the following line of code:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] <- mtcars$mpg[mtcars$cyl>6]^2
```

The same expression, `mtcars$mpg[mtcars$cyl>6]`, is written twice, making this code rather long and cumbersome, even though we're just squaring the expression. 

This R package solves the above laid-out problem by implementing a generalized in-place (mathematical) modifier, through the `x %:=% f` operator.

With `tinyoperations` one can now make this more compact (more "tiny", if you will) as follows:

```{r eval=FALSE}
mtcars$mpg[mtcars$cyl>6] %:=% \(x)x^2
```



&nbsp;

