---
title: "Miscellaneous functionality"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Miscellaneous functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinyoperations)
```


&nbsp;



# source_selection

The `source_selection()` function is the same as base R's `source()` function, except that it allows only placing the selected objects and functions into the current environment, instead of all objects.

The objects to be selected can be specified using any combination of the following:

 - by supplying a character vector of exact object names to the `select` argument.
 - by supplying a character vector of `regex` patterns to the `regex` argument.
 - by supplying a character vector of `fixed` patterns to the `fixed` argument.

The `regex` and `fixed` arguments are especially handy when sourcing S3 methods. For example, to expose the following methods to the current environment,
`mymethod.numeric()` and `mymethod.character()` from generic `mymethod()`,
one could specify `regex = "^mymethod"`.

Example:

```{r}
exprs <- expression({
  helloworld = function()print("helloworld")
  goodbyeworld <- function() print("goodbye world")
  `%s+test%` <- function(x,y) stringi::`%s+%`(x,y)
  `%s*test%` <- function(x,y) stringi::`%s*%`(x,y)
  mymethod <- function(x) UseMethod("mymethod", x)
  mymethod.numeric <- function(x)x * 2
  mymethod.character <- function(x)chartr(x, old = "a-zA-Z", new = "A-Za-z")
})

temp.fun <- function(){
  source_selection(list(exprs=exprs), regex= "^mymethod", fixed = c("%", ":="))
  ls() # list all objects residing within the function definition
}
temp.fun()

temp.fun <- function(){
  source_selection(list(exprs=exprs), select = c("helloworld", "goodbyeworld"))
  ls() # list all objects residing within the function definition
}
temp.fun()

```


&nbsp;


# Matrix re-ordering operators

The `tinyoperations` R package adds 2 additional matrix operators:

 - The `x %row~% mat` operator  re-orders the elements of every row, each row ordered independently from the other rows, of matrix `x`, by the ordering ranks given in matrix `mat`.
 - The `x %col~% mat` operator re-orders the elements of every column, each column ordered independently from the other columns, of matrix `x`, by the ordering ranks given in matrix `mat`.
 
If matrix `x` is a numeric matrix, and one wants to numerically sort the elements of every row or column, `x %row~% x` or `x %col~% x` would suffice, respectively.

If matrix `x` is not numeric, sorting using `x %row~% x` and `x %col~% x` is still possible, but probably not the best option. In the non-numeric case, providing a matrix of ordering ranks would probably be faster and give more accurate ordering.

If `mat` is a matrix of non-repeating random integers (i.e. `sample(1:length(x), replace=FALSE)`), `x %row~% mat` will randomly shuffle the elements of every row, where the shuffling order in each row is independent from the shuffling order in the other rows. Similarly, `x %col~% mat` will randomly shuffle the elements of every column, where the shuffling order in each column is independent from the shuffling order in the other columns.

Re-ordering/sorting every row/column of a matrix with these operators is generally faster than doing so through loops or apply-like functions.

&nbsp;

Examples with a numeric matrix:

```{r}
x <- matrix(sample(1:25), nrow=5)
print(x)
x %row~% x # sort elements of every row
x %row~% -x # reverse-sort elements of every row
x %col~% x # sort elements of every column
x %col~% -x # reverse-sort elements of every column

x <- matrix(sample(1:25), nrow=5)
print(x)
rand <- sample(1:length(x)) |> matrix(ncol=ncol(x)) # matrix of random integers
x %row~% rand # random shuffle every row independent of other rows
x %col~% rand # random shuffle every column independent of other columns
```


Examples with a character matrix:

```{r}
x <- matrix(sample(letters, 25), nrow=5)
print(x)
mat <- stringi::stri_rank(as.vector(x)) |>
  matrix(ncol=ncol(x)) # matrix of ordering ranks
x %row~% mat # sort elements of every row
x %row~% -mat # reverse-sort elements of every row
x %col~% mat # sort elements of every column
x %col~% -mat # reverse-sort elements of every column

x <- matrix(sample(letters, 25), nrow=5)
print(x)
rand <- sample(1:length(x)) |> matrix(ncol=ncol(x)) # matrix of random integers
x %row~% rand # random shuffle every row independent of other rows
x %col~% rand # random shuffle every column independent of other columns
```



&nbsp;

# Additional logic operators

The tinyoperations package adds a few basic logic operators:

 - `%xor%`: Exclusive OR
 - `%n&%`: NOT AND (i.e. `(!x) & (!y)`). Note that if either `x` or `y` is `NA`, `%n&%` will also give `NA` (unlike `(!x) & (!y)`, which would give `FALSE`.)
 - `%?=%`: checks if both `x` and `y` are unknown or unreal (NA, NaN, Inf, -Inf)
 - `%out%`: the opposite of `%in%` (i.e. `!x %in% y`)

Here are some examples:

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, NA, NaN, Inf, -Inf, TRUE, FALSE)
y <- c(FALSE, TRUE, TRUE, FALSE, rep(NA, 6))
outcome <- data.frame(
  x=x, y=y,
  "x %xor% y"=x %xor% y, "x %n&% y" = x %n&% y, "x %?=% y" = x %?=% y,
  check.names = FALSE
)
knitr::kable(outcome)

1:3 %out% 1:10
1:10 %out% 1:3
```



Numbers can have many different sub-types whilst still being `numeric`. The `n %=numtype% numtype` operator will check for every value of numeric vector `n` if it can be considered a number belonging to type `numtype`. The following values for `numtype` are allowed:

 - "~0": zero, or else a number whose absolute value is smaller than the Machine tolerance (`sqrt(.Machine$double.eps)`);
 - "B": binary numbers (0 or 1);
 - "prop": proportions;
 - "I": Integers;
 - "odd": odd integers;
 - "even": even integers
 - "R": Real numbers;
 - "unreal": infinity, NA, or NaN;

The string counterpart for `%=numtype%` is `s %=strtype% strtype`, which checks for every value of character vector `s` if it can seen as a certain `strtype`. The following values for `strtype` are allowed:

 - "empty": checks if the string only consists of empty spaces.
 - "unreal": checks if the string is NA, or if it has literal string "NA", "NaN" or "Inf", regardless if it has leading or trailing spaces.
 - "numeric": checks if the string can be converted to a number, disregarding leading and trailing spaces. I.e. the string "5.0" can be converted to the the actual number 5.0.
 - "special": checks if the string consists of only special characters.

Here are some examples:

```{r}

1e-20 %=numtype% "~0"
n <- c(0:5, 0:-5, 0.1, -0.1, 0, 1, Inf, -Inf, NA, NaN)
n[n %=numtype% "B"]
n[n %=numtype% "prop"]
n[n %=numtype% "B"]
n[n %=numtype% "I"]
n[n %=numtype% "odd"]
n[n %=numtype% "even"]
n[n %=numtype% "R"]
n[n %=numtype% "unreal"]

s <- c(" AbcZ123 ", " abc ", " 1.3 ", " !#$%^&*() ", "  ", "  NA  ", "  NaN  ", " Inf ")
s[s %=strtype% "empty"]
s[s %=strtype% "unreal"]
s[s %=strtype% "numeric"]
s[s %=strtype% "special"]
```


&nbsp;


