---
title: "Relations to other packages"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Relations to other packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
```

&nbsp;

# Introduction

This article lays out how the `tinycodet` R package relates to other R packages, and compatibility with other R packages, etc.

&nbsp;

# Compatibility with other R packages

## Name conflicts

I made sure not to name any of the operators in `tinycodet` the same as the operators in the popular operator-related packages: [magrittr](https://github.com/tidyverse/magrittr) and [zeallot](https://github.com/r-lib/zeallot). 

One may have noticed that some of the functionality in this R packages has been inspired by other programming languages, in particular by Python (please, refrain from taking out your torches and pitchforks!). This is not the first R package to take inspiration from other programming languages like Python. Therefore, one might find some overlapping function and operator names in other packages that do something similar; so be mindful of that.

&nbsp;

## Other import systems

The [import](https://github.com/rticulate/import) R package provides somewhat similar capabilities to the `tinycodet` import management system, but they are still quite different. The `tinycodet`' import system focuses more on importing a main package + its re-exports + (reverse) dependencies under a alias, and exposing infix operators from a package to the current environment. The [import](https://github.com/rticulate/import) package does not really provide this functionality directly (as far as I know). Strictly speaking there is no incompatibility between `tinycodet` and [import](https://github.com/rticulate/import). You can safely use both implementations of the import system, if you wish.

&nbsp;

There is also the [box](https://github.com/klmr/box) package. Where [import](https://github.com/rticulate/import) is primarily for importing packages, [box](https://github.com/klmr/box) is primarily for sourcing user-defined scripts as modules, though it can also import packages. Like [import](https://github.com/rticulate/import), [box](https://github.com/klmr/box) does not support importing multiple packages (or modules) under the same alias. It is compatible with `tinycodet`'s package import system; one can use [box](https://github.com/klmr/box) for handling user-defined scripts, and `tinycodet` for handling packages.

&nbsp;

## Renv

When using the [renv](https://github.com/rstudio/renv) R package, note that it only registers packages that are attached using plain `library()` or `require()` calls. Anything different from that, whether it be loading (for example `::` or `loadNamespace()`), importing, or even things like `for(... in ...)library(...)` or `if(...)library(...)`, will by default NOT be understood by [renv](https://github.com/rstudio/renv). Therefore, if using [renv](https://github.com/rstudio/renv), please make sure to set the following:

```{r eval=FALSE}
renv::settings$snapshot.type("all")
```

This will make sure that all packages installed in your project library, regardless of how they are loaded,  will all be registered by [renv](https://github.com/rstudio/renv). This makes [renv](https://github.com/rstudio/renv) compatible with calls like  `import_as` from `tinycodet`, and things like `for(... in ...)library(...)` or `if(...)library(...)`.

&nbsp;


# Recommended R packages

[stringi](https://github.com/gagolews/stringi) is of course required for this packages.

Besides that, I highly recommend the [fastverse](https://github.com/fastverse/fastverse) set of R packages (https://github.com/fastverse/fastverse), which are a set of R packages for (mostly) data wrangling, focused on high speed, better memory management, and minimal dependencies.

For quick 'n easy back-tracing of errors, I recommend the [rlang](https://github.com/r-lib/rlang/tree/main/R) R package.




&nbsp;
