---
title: "Benchmarks"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tinycodet)
loadNamespace("bench")
loadNamespace("ggplot2")
```

&nbsp;

# Introduction

Although the functions in this R package were written R, the functions - especially the string related ones - have been somewhat aggressively optimized. The string related functions, for example, are about in the same order of magnitude in terms of speed as the `stringi` functions they call.


Here some speed comparisons are given, using the 'bench' package.

&nbsp;

# stri_locate_ith

```{r eval = FALSE}
n <- 1e5
x <- rep(paste0(1:50, collapse = ""), n)
p <- "\\d"
i <- sample(c(-50:-1, 1:50), replace=TRUE, size = n)
locate_stringi <- function(...) {
  stringi::stri_locate_all(...)
  stringi::stri_count(...)
}
bm.stri_locate_ith <- bench::mark(
  "stri_locate_ith" = { stri_locate_ith_regex(x, p, i) },
  "stringi::(stri_locate_all + stri_count)" = { locate_stringi(str=x, regex = p) },
  min_iterations = 500,
  check = FALSE,
  filter_gc = FALSE
)
ggplot2::autoplot(bm.stri_locate_ith)
```

```{r echo = FALSE, fig.width = 8}
load("bm.stri_locate_ith.RData")
ggplot2::autoplot(bm.stri_locate_ith)
```


&nbsp;

# strcut

```{r eval = FALSE}
n <- 1e5
x <- rep("hello", n)
i <- sample(1:3, n, replace = TRUE)
loc <- stri_locate_ith(x, i=i, regex="a|e|i|o|u")
bm.strcut <- bench::mark(
  "strcut_loc" = { strcut_loc(x, loc) },
  "strcut_brk" = { strcut_brk(x, type = "", tolist = TRUE) },
  "stringi::stri_split_boundaries" = {
    stringi::stri_split_boundaries(x, type="character")
  },
  min_iterations = 500,
  check = FALSE,
  filter_gc = FALSE
)
ggplot2::autoplot(bm.strcut)
```

```{r echo = FALSE, fig.width=8}
load("bm.strcut.RData")
ggplot2::autoplot(bm.strcut)
```

&nbsp;

# Row/column-wise re-ordering

```{r eval = FALSE}
mat <- matrix(1:(1e3)^2, ncol = 1e3)
bm.matorder <- bench::mark(
  tinycodet = {mat %row~% mat; mat %col~% mat},
  Rfast = {Rfast::rowSort(mat); Rfast::colSort(mat)},
  min_iterations = 1000
)
ggplot2::autoplot(bm.matorder)
```

```{r echo = FALSE, fig.width=8}
load("bm.matorder.RData")
ggplot2::autoplot(bm.matorder)
```

