---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(badger)
library(knitr)
library(tinycodet)
set.seed(1)
```


# Introduction

The `tinycodet` R-package adds some functions to help in your coding etiquette. It primarily focuses on 4 aspects:

 1) Safer decimal (in)equality testing, safer atomic typecasting, standard-evaluated alternatives to `with()` and `aes()`, and other functions for safer coding.
 2) A new package import system, that attempts to combine the benefits of using a package without attaching, with the benefits of attaching a package.
 3) Extending the string manipulation capabilities of the `stringi` R package.
 4) Reducing repetitive code.

The `tinycodet` R-package has only one dependency, namely `stringi`. Most functions in this R-package are fully vectorized and optimized, and have been well documented.

Here I'll give a quick glimpse of what is possible in this R package.

&nbsp;


# Safer functionality

'tinycodet' adds some functions to help in coding more safely:

 - In base R, `(0.1*3) == 0.3` gives `FALSE`, due to the way decimal numbers are stored in programming languages like R and Python.  `tinycodet` adds safer truth testing operators, that give correct results.
 - The atomic type casting functions (i.e. `as.character()`, `as.logical()`, etc.) strip away attributes. `tinycodet` adds safer type casting functions that do not strip away attributes - except the class attribute.
 - `with_pro()` and `aes_pro()` are standard-evaluated alternatives to `base::with()` and `ggplot2::aes()`. These use formulas as input (see also the `form()` function for constructing formula's without memory leakage).
 - One can re-assign the values `T` and `F`. One can even run `T <- FALSE` and `F <- TRUE`!. `tinycodet` adds the `lock_TF()` function that forces `T` to stay `TRUE` and `F` to stay `FALSE`.

One example with the `%d==%` operator:

```{r}
x <- c(0.3, 0.6, 0.7)
y <- c(0.1*3, 0.1*6, 0.1*7)
print(x); print(y)

x == y # gives FALSE, but should be TRUE

x %d==% y # here it's done correctly
```


&nbsp;

# New import system

One can use a package without attaching the package (for example using `::`), or one can attach a package (for example using `library()` or `require()`). The advantages and disadvantages of using without attaching a package versus attaching a package - at least those relevant for now - can be compactly presented in the following table:

```{r echo=FALSE, results='markup'}
X <- rbind(
  # c(aspect, alias, attach)
  c("prevent masking functions from other packages", "Yes (+)", "No (-)"),
  c("prevent masking core R functions", "Yes (+)", "No (-)"),
  c("clarify which function came from which package", "Yes (+)", "No (-)"),
  c("place/expose functions only in current environment instead of globally", "Yes (+)", "No (-)"),
  c("prevent namespace pollution", "Yes (+)", "No (-)"),
  c(
  "minimize typing - especially for infix operators <br> (i.e. typing ``package::`%op%`(x, y)`` instead of `x %op% y` is cumbersome)", "No (-)", "Yes (+)"),
  c("use multiple related packages, <br> without constantly switching between package prefixes", "No (-)", "Yes (+)"),
  c("NOTE: + = advantage, - = disadvantage", "", "")
)
colnames(X) <- c("aspect", "::", "attach")
X <- as.data.frame(X)
rownames(X) <- c(1:7, "")
knitr::kable(X, "html", row.names = TRUE, escape = FALSE)
```


What `tinycodet` attempts to do with its import system, is to somewhat find the best of both worlds. It does this by introducing the following functions:

 - `import_as()`: Import a main package, and optionally its re-exports + its dependencies + its extensions, under a single alias. This essentially combines the attaching advantage of using multiple related packages (row 7 on the table above), whilst keeping most advantages of using without attaching a package.
 - `import_inops()`: Expose infix operators from a package or an alias object to the current environment. This gains the attaching advantage of less typing (row 6 in table above), whilst simultaneously avoiding the disadvantage of attaching functions from a package globally (row 4).
 - `import_data()`: Directly return a data set from a package, to allow straight-forward assignment.

Here is an example using `tinycodet's` new import system; note that the following code is run without attaching a single R package (besides `tinycodet` itself of course):

```{r}
# importing "tidytable" + "data.table" under alias "tdt.":
import_as( 
  ~ tdt., "tidytable", dependencies = "data.table"
)

# exposing operators from `magrrittr` to current environment:
import_inops("magrittr")

# directly assigning the "starwars" dataset to object "d":
d <- import_data("dplyr", "starwars") 

# see it in action:
d %>% tdt.$filter(species == "Droid") %>%
  tdt.$select(name, tdt.$ends_with("color"))

```


&nbsp;

# Extending the string manipulation capabilities of stringi

```{r}

# character vector:
x <- c("3rd 1st 2nd", "5th 4th 6th")
print(x)

# detect if there are digits:
x %s{}% "\\d"

# find second last digit:
loc <- stri_locate_ith(x, i = -2, regex = "\\d")
stringi::stri_sub(x, from = loc)

# cut x into matrix of individual words:
mat <- strcut_brk(x, "word")

# sort rows of matrix using the fast %row~% operator:
rank <- stringi::stri_rank(as.vector(mat)) |> matrix(ncol = ncol(mat))
sorted <- mat %row~% rank
sorted[is.na(sorted)] <- ""

# join elements of every row into a single character vector:
stri_c_mat(sorted, margin = 1, sep = " ")


```



&nbsp;

# Reduce repetitive code

```{r eval=FALSE}
# in base R:
ifelse( # repetitive, and gives unnecessary warning
  is.na(object>0), -Inf,
  ifelse(
    object>0,  log(object), object^2
  )
)
mtcars$mpg[mtcars$cyl>6] <- (mtcars$mpg[mtcars$cyl>6])^2 # long

# with tinycodet:
object |> transform_if(\(x)x>0, log, \(x)x^2, \(x) -Inf) # compact & no warning
mtcars$mpg[mtcars$cyl>6] %:=% \(x)x^2 # short
```

&nbsp;

# The Articles

If you're still interested, I invite you to read the articles on the website (https://tony-aw.github.io/tinycodet/), and perhaps try out the package yourself.

The following articles are currently present:

 - [Safer functionality](https://tony-aw.github.io/tinycodet/articles/a_safer.html): Describes the safer decimal (in)equality testing operators, safer atomic conversion functions, and other functions for safer coding.
 - [Import system - main functions](https://tony-aw.github.io/tinycodet/articles/b_import_main.html): Description of the main functions of the package import system introduced by `tinycodet`.
 - [Import system - additional details](https://tony-aw.github.io/tinycodet/articles/c_import_additional.html): Additional important details on the tinycodet import system.
 - [String related functions](https://tony-aw.github.io/tinycodet/articles/d_strings_functions.html): Describes the `tinycodet` functions that extend the string manipulation capabilities of `stringi`.
 - [String related infix operators](https://tony-aw.github.io/tinycodet/articles/e_strings_inops.html): Describes the `tinycodet` infix operators that extend the string manipulation capabilities of `stringi`.
 - [Don't Repeat Yourself](https://tony-aw.github.io/tinycodet/articles/f_DRY.html): Describes the `tinycodet` functions that help reduce repetitions in your code.
 - [Miscellaneous functionality](https://tony-aw.github.io/tinycodet/articles/g_misc.html): Various other functions and operators that `tinycodet` introduces.
 - [Relations to other R packages](https://tony-aw.github.io/tinycodet/articles/h_otherpkgs.html): Describes how `tinycodet` relates to other R packages, mostly regarding compatibility.
 
For a complete list of functions introduced by `tinycodet`, please see the [References page](https://tony-aw.github.io/tinycodet/reference/index.html).


```{r echo=FALSE}
rm(list=ls())
```

&nbsp;
